# Household Hunger Scale 

## References / Resources

1. https://www.fantaproject.org/monitoring-and-evaluation/household-hunger-scale-hhs


## Standardized Questionnaire

Variable Name | Question Label | Answer Choices
---- | ------------- |---- 
HHhSNoFood_YN | In the past [4 weeks/30 days], was there ever no food to eat of any kind in your house because of lack of resources to get food? | 1) Yes 0) No
HHhSNoFood_FR | How often did this happen in the past [4 weeks/30 days]? | 1) Rarely (1–2 times) 2) Sometimes (3–10 times)  3) Often (more than 10 times)
HHhSBedHung_YN | In the past [4 weeks/30 days], did you or any household member go to sleep at night hungry because there was not enough food? | 1) Yes 0) No
HHhSBedHung_FR  | How often did this happen in the past [4 weeks/30 days]? | 1) Rarely (1–2 times) 2) Sometimes (3–10 times)  3) Often (more than 10 times)
HHhSNotEat_YN | In the past [4 weeks/30 days], did you or any household member go to sleep at night hungry because there was not enough food? | 1) Yes 0) No
HHhSNotEat_FR | How often did this happen in the past [4 weeks/30 days]? | 1) Rarely (1–2 times) 2) Sometimes (3–10 times)  3) Often (more than 10 times)

### Paper Version of Questionnaire

Here is the standardized module in word format:
[RBDstandardized_questionnaireHHS](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/questionnaires/RBDstandardized_questionnaireHHS.xlsx) 

### Electronic Version of Questionnaire

Here is the standardized module in xlsform format:
[RBDstandardized_questionnaireHHS](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/questionnaires/RBDstandardized_questionnaireHHS.docx) 


## Calculation of Household Hunger Scale Indicators & Cadre Harmonise Phasing

### Example data set

Here is the example data set:
[dataHHSEng](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/example_datasets/dataHHSEng.sav) 

### SPSS Syntax

```{r, eval = F}

GET  FILE='dataHHSEng.sav'.
DATASET NAME DataSet1 WINDOW=FRONT.

** recode the frequency questions to scores

Recode HHhSNoFood_FR HHhSBedHung_FR HHhSNotEat_FR (1 = 1) (2=1) (3=2) (ELSE=0) INTO HHhSNoFood_FR_r HHhSBedHung_FR_r HHhSNotEat_FR_r.

Variable labels HHhSNoFood_FR_r "In the past [4 weeks/30 days], was there ever no food to eat of any kind in your house because of lack of resources to get food? - recoded"
HHhSBedHung_FR_r "In the past [4 weeks/30 days], did you or any household member go to sleep at night hungry because there was not enough food? - recoded"
HHhSNotEat_FR_r "In the past [4 weeks/30 days], did you or any household member go to sleep at night hungry because there was not enough food? - recoded".

** sum the recoded questions to calculate the HHS

Compute HHhS = HHhSNoFood_FR_r + HHhSBedHung_FR_r + HHhSNotEat_FR_r.
variable labels HHhS "Household Hunger Scale".

** each household should have an HHhS score between 0 - 6.

FREQUENCIES VARIABLES = HHhS
/STATISTICS=MEAN MEDIAN MINIMUM MAXIMUM
/ORDER=ANALYSIS.

** Create Categorical HHhS

RECODE HHhS (0 thru 1=1) (2 thru 3=2) (4 thru Highest=3) INTO HHhSCat.
variable labels HHhSCat "Household Hunger Score Categories".
value labels HHhSCat 
1 `No or little hunger in the household`
2 `Moderate hunger in the household`
3 `Severe hunger in the household`.

** Create HHhS Cadre Harmonise Categories

RECODE HHS (0=1) (1=2) (2 thru 3=3) (4=4) (5=5) INTO HHhS_CH.
variable labels HHhS_CH "Household Hunger Score Categories - Cadre Harmonise".
value labels HHhS_CH
1 `Phase1`
2 `Phase2`
3 `Phase3`
4 `Phase4`
5 `Phase5`.

** Generate table of proportion of households in CH HHS phases by Adm1 and Adm2 using weights

WEIGHT BY WeightHH.

CROSSTABS
  /TABLES=ADMIN2Name BY HHhS_CH BY ADMIN1Name
  /FORMAT=AVALUE TABLES
  /CELLS=ROW 
  /COUNT ROUND CELL.

```

Here is the SPSS syntax file: 

[RBDstandardized_SPSSsyntaxHHS](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/syntax/RBDstandardized_SPSSsyntaxHHS.sps) 

### R Syntax

```{r, eval = T, message = F}
library(haven)
library(labelled)
library(tidyverse)

#import dataset
dataHHSEng <- read_sav("example_datasets/dataHHSEng.sav")

#Calculate HHS 
dataHHSEng <- to_factor(dataHHSEng)

#Recode HHS questions into new variables with score 
dataHHSEng <- dataHHSEng %>% mutate(HHhSNoFood_FR_r = case_when(
                                      HHhSNoFood_FR == "Rarely (1–2 times)" ~ 1,
                                      HHhSNoFood_FR == "Sometimes (3–10 times)" ~ 1,
                                      HHhSNoFood_FR == "Often (more than 10 times)" ~ 2,
                                      TRUE ~ 0),
                                    HHhSBedHung_FR_r = case_when(
                                      HHhSBedHung_FR == "Rarely (1–2 times)" ~ 1,
                                      HHhSBedHung_FR == "Sometimes (3–10 times)" ~ 1,
                                      HHhSBedHung_FR == "Often (more than 10 times)" ~ 2,
                                      TRUE ~ 0),
                                    HHhSNotEat_FR_r = case_when(
                                      HHhSNotEat_FR == "Rarely (1–2 times)" ~ 1,
                                      HHhSNotEat_FR == "Sometimes (3–10 times)" ~ 1,
                                      HHhSNotEat_FR == "Often (more than 10 times)" ~ 2,
                                      TRUE ~ 0))
# Calculate HHhS score
dataHHSEng <- dataHHSEng %>% mutate(HHhS = HHhSNoFood_FR_r + HHhSBedHung_FR_r + HHhSNotEat_FR_r)
var_label(dataHHSEng$HHhS) <- "Household Hunger Scale"

#each household should have an HHS score between 0 - 6
summary(dataHHSEng$HHhS)

# Create Categorical HHS
dataHHSEng <- dataHHSEng %>% mutate(HHhSCat = case_when(
    HHhS %in% c(0,1) ~ "No or little hunger in the household",
    HHhS %in% c(2,3) ~ "Moderate hunger in the household",
    HHhS >= 4 ~ "Severe hunger in the household" 
    ))
var_label(dataHHSEng$HHhSCat) <- "Household Hunger Score Categories"

#Convert HH Scores to CH phases
dataHHSEng <- dataHHSEng %>% mutate(HHhS_CH = case_when(
    HHhS == 0 ~ "Phase1",
    HHhS == 1 ~ "Phase2",
    HHhS %in% c(2,3) ~ "Phase3",
    HHhS == 4 ~ "Phase4",  
    HHhS >= 5 ~ "Phase5"))
var_label(dataHHSEng$HHhS_CH) <- "Household Hunger Score Categories - Cadre Harmonise"

#Generate table of proportion of households in CH HHS phases by Adm1 and Adm2 using weights
CH_HHhS_table_wide <- dataHHSEng %>% group_by(ADMIN1Name, ADMIN2Name) %>%
  drop_na(HHhS_CH) %>%
  count(HHhS_CH, wt = WeightHH) %>%
  mutate(perc = 100 * n / sum(n)) %>%
  ungroup() %>% select(-n) %>%
  spread(key = HHhS_CH, value = perc) %>% replace(., is.na(.), 0)  %>% mutate_if(is.numeric, round, 1)
```

Here is the R syntax file: 

[RBDstandardized_RsyntaxHHS](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/syntax/RBDstandardized_RsyntaxHHS.R) 


## Visualisation & Data Quality Monitoring of HHS Indicator

1. Compare HHS CH phasing by admin1 and admin2 areas

```{r, eval = T, message = F}
#Generate longtable of proportion of households in CH HHS phases by Adm1 and Adm2 not using weights
CH_HHhS_table_long <- dataHHSEng %>% group_by(ADMIN1Name, ADMIN2Name) %>%
  drop_na(HHhS_CH) %>%
  count(HHhS_CH) 

#make graph
CH_HHhS_admin2_bargraph <- CH_HHhS_table_long %>% ggplot(aes(fill=HHhS_CH, y=n, x=ADMIN2Name)) +geom_bar(position="fill", stat="identity") +facet_wrap(. ~ ADMIN1Name, scales = "free") +scale_fill_manual(values=c("#c6ffc7","#ffe718","#e88400","#e02d00", "#5e0803")) +theme_minimal() +theme(axis.text.x = element_text(angle = 90)) +scale_y_continuous(labels = scales::percent)+ theme(axis.title.x = element_blank(), axis.title.y = element_blank()) 

#add title
CH_HHhS_admin2_bargraph <- CH_HHhS_admin2_bargraph +ggtitle("CH Phasing of HHS Groups of Households by Admin2 Area") +theme(plot.title = element_text(hjust = 0.5))
```


```{r, eval = T, message = F}
library(plotly)

ggplotly(CH_HHhS_admin2_bargraph)
```


2. Examine HHS CH phasing by admin2 areas and enumerators 

```{r, eval = T, message = F}
#Generate longtable of proportion of households in CH HHS phases by Adm1 and Adm2 and Operator - not using survey weights 
CH_HHhS_enumerators_tablelong <- dataHHSEng  %>% 
  group_by(ADMIN2Name, enum_cod) %>%
  count(HHhS_CH) 


#make graph
CH_HHhS_admin2enumerators_bargraph <- CH_HHhS_enumerators_tablelong %>% ggplot(aes(fill=HHhS_CH, y=n, x=enum_cod)) +geom_bar(position="fill", stat="identity") +facet_wrap(. ~ ADMIN2Name, scales = "free") +scale_fill_manual(values=c("#c6ffc7","#ffe718","#e88400","#e02d00", "#5e0803")) +theme_void()

#add title
CH_HHhS_admin2enumerators_bargraph  <- CH_HHhS_admin2enumerators_bargraph +ggtitle("CH Phasing of HHS Groups of Households by Admin2 Area and Enumerator") +theme(plot.title = element_text(hjust = 0.5))

```




```{r, eval = T, message = F}
library(plotly)

ggplotly(CH_HHhS_admin2enumerators_bargraph)
```








