# Household Hunger Scale {#HHS}

## Standardized Questionnaire

Variable Name | Question Label | Answer Choices
---- | ------------- |---- 
HHhSNoFood_YN | In the past [4 weeks/30 days], was there ever no food to eat of any kind in your house because of lack of resources to get food? | 1) Yes 0) No
HHhSNoFood_FR | How often did this happen in the past [4 weeks/30 days]? | 1) Rarely (1–2 times) 2) Sometimes (3–10 times)  3) Often (more than 10 times)
HHhSBedHung_YN | In the past [4 weeks/30 days], did you or any household member go to sleep at night hungry because there was not enough food? | 1) Yes 0) No
HHhSBedHung_FR  | How often did this happen in the past [4 weeks/30 days]? | 1) Rarely (1–2 times) 2) Sometimes (3–10 times)  3) Often (more than 10 times)
HHhSNotEat_YN | In the past [4 weeks/30 days], did you or any household member go to sleep at night hungry because there was not enough food? | 1) Yes 0) No
HHhSNotEat_FR | How often did this happen in the past [4 weeks/30 days]? | 1) Rarely (1–2 times) 2) Sometimes (3–10 times)  3) Often (more than 10 times)

Here is the standardized module in xlsform:

[RBDstandardized_questionnaireHHS](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/questionnaires/RBDstandardized_questionnaireHHS.xlsx) 

## Calculation of HHS Indicator & CH 

### Example data set

Here is the example data set:
[dataHHSEng](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/example_datasets/dataHHSEng.sav) 

### SPSS Syntax

```{r, eval = F}

to be added

```

Here is the SPSS syntax file: 
[RBDstandardized_spsssyntaxrCSI](https://www.dropbox.com/s/gsj4tvadww9v3q4/RBDstandardized_questionnairerCSI.xlsx?dl=0) 


### R Syntax

```{r, eval = F}
library(haven)
library(labelled)
library(tidyverse)

#import dataset
dataHHSEng <- read_sav("dataHHSEng.sav")

#Calculate HHS 
dataHHSEng <- to_factor(dataHHSEng)

#Recode HHS questions into new variables with score 
dataHHSEng <- dataHHSEng %>% mutate(HHhSNoFood_FR_r = case_when(
                                      HHhSNoFood_FR == "Rarely (1–2 times)" ~ 1,
                                      HHhSNoFood_FR == "Sometimes (3–10 times)" ~ 1,
                                      HHhSNoFood_FR == "Often (more than 10 times)" ~ 2,
                                      TRUE ~ 0),
                                    HHhSBedHung_FR_r = case_when(
                                      HHhSBedHung_FR == "Rarely (1–2 times)" ~ 1,
                                      HHhSBedHung_FR == "Sometimes (3–10 times)" ~ 1,
                                      HHhSBedHung_FR == "Often (more than 10 times)" ~ 2,
                                      TRUE ~ 0),
                                    HHhSNotEat_FR_r = case_when(
                                      HHhSNotEat_FR == "Rarely (1–2 times)" ~ 1,
                                      HHhSNotEat_FR == "Sometimes (3–10 times)" ~ 1,
                                      HHhSNotEat_FR == "Often (more than 10 times)" ~ 2,
                                      TRUE ~ 0))
#Calculate HHS
dataHHSEng <- dataHHSEng %>% mutate(HHhS = HHhSNoFood_FR_r + HHhSBedHung_FR_r + HHhSNotEat_FR_r)

#Convert HH Scores to CH phases
dataHHSEng <- dataHHSEng %>% mutate(CH_HHS = case_when(
    HHhS == 0 ~ "Phase1",
    HHhS == 1 ~ "Phase2",
    HHhS %in% c(2,3,4) ~ "Phase3",
    HHhS == 4 ~ "Phase4",  
    HHhS >= 5 ~ "Phase5"))

#Generate table of proportion of households in CH HHS phases by Adm1 and Adm2 using weights
CH_HHS_table_wide <- dataHHSEng %>% group_by(ADMIN1Name, ADMIN2Name) %>%
  drop_na(CH_HHS) %>%
  count(CH_HHS, wt = hh_weight) %>%
  mutate(perc = 100 * n / sum(n)) %>%
  ungroup() %>% select(-n) %>%
  spread(key = CH_HHS, value = perc) %>% replace(., is.na(.), 0)  %>% mutate_if(is.numeric, round, 1)
```


Here is the R syntax file: 
[RBDstandardized_RsyntaxHHS](https://www.dropbox.com/s/xnzi9vfsapakhvo/RBDstandardized_RsyntaxHHS.R?dl=0) 
