# reduced Coping Strategy Index {#rCSI}

## Standardized Questionnaire

Variable Name | Question Label 
---- | ------------- 
rCSILessQlty | In the past 7 days, how many days has your household had to: *Rely on less preferred and less expensive food* because you did not have enough food or money to buy food?
rCSIBorrow | In the past 7 days, how many days has your household had to: *Borrow food or rely on help from a relative or friend* because you did not have enough food or money to buy food?
rCSIMealSize | In the past 7 days, how many days has your household had to: *Limit portion size of meals* because you did not have enough food or money to buy food?
rCSIMealAdult  | In the past 7 days, how many days has your household had to: *Restrict consumption by adults in order for small children to eat* because you did not have enough food or money to buy food?
rCSIMealNb | In the past 7 days, how many days has your household had to: *Reduce the number of meals eaten per day* because you did not have enough food or money to buy food?

Here is the standardized module in xlsform:

[RBDstandardized_questionnaireCSI](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/questionnaires/RBDstandardized_questionnairerCSI.xlsx) 


## Calculation of rCSI Indicator & CH 

### Example data set

Here is the example data set:
[datarCSIEng](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/example_datasets/dataHHSEng.sav) 

### SPSS Syntax

```{r, eval = F}

to be added

```

Here is the SPSS syntax file: 

not done yet

### R Syntax

```{r, eval = F}
library(haven)
library(labelled)
library(tidyverse)

#import dataset
datarCSIEng <- read_sav("datarCSIEng.sav")

#Calculate HHS 
datarCSIEng <- to_factor(datarCSIEng)

#calculate rCSI Score
datarCSIEng <- datarCSIEng %>% mutate(rCSI = rCSILessQlty  + (2 * rCSIBorrow) + rCSIMealSize + (3 * rCSIMealAdult) + rCSIMealNb)

#convert rCSI Score into thresholds 
datarCSIEng <- datarCSIEng %>% mutate(CH_rCSI = case_when(
  rCSI <= 3 ~ "Phase1", 
  between(rCSI,4,18) ~ "Phase2",       
  rCSI >= 19 ~ "Phase3"))

#Generate table of proportion of households in CH rCSI phases by Adm1 and Adm2 using weights
CH_rCSI_table_wide <- datarCSIEng %>% group_by(ADMIN1Name, ADMIN2Name) %>%
  drop_na(CH_rCSI) %>%
  count(CH_rCSI, wt = hh_weight) %>%
  mutate(perc = 100 * n / sum(n)) %>%
  ungroup() %>% select(-n) %>%
  spread(key = CH_rCSI, value = perc) %>% replace(., is.na(.), 0)  %>% mutate_if(is.numeric, round, 1)
```



Here is the R syntax file: 

[RBDstandardized_RsyntaxrCSI](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/syntax/RBDstandardized_RsyntaxrCSI.R) 


## References / Resources

