# reduced Coping Strategy Index 

## References / Resources

1. https://documents.wfp.org/stellent/groups/public/documents/manual_guide_proced/wfp211058.pdf

2. https://documents.wfp.org/stellent/groups/public/documents/manual_guide_proced/wfp271449.pdf

## Standardized Questionnaire

Variable Name | Question Label 
---- | ------------- 
rCSILessQlty | In the past 7 days, how many days has your household had to: *Rely on less preferred and less expensive food* because you did not have enough food or money to buy food?
rCSIBorrow | In the past 7 days, how many days has your household had to: *Borrow food or rely on help from a relative or friend* because you did not have enough food or money to buy food?
rCSIMealSize | In the past 7 days, how many days has your household had to: *Limit portion size of meals* because you did not have enough food or money to buy food?
rCSIMealAdult  | In the past 7 days, how many days has your household had to: *Restrict consumption by adults in order for small children to eat* because you did not have enough food or money to buy food?
rCSIMealNb | In the past 7 days, how many days has your household had to: *Reduce the number of meals eaten per day* because you did not have enough food or money to buy food?


### Paper Version of Questionnaire

Here is the standardized module in word format:
[RBDstandardized_questionnairerCSI](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/questionnaires/RBDstandardized_questionnairerCSI.docx) 

### Electronic Version of Questionnaire

Here is the standardized module in xlsform:
[RBDstandardized_questionnairerCSI](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/questionnaires/RBDstandardized_questionnairerCSI.xlsx) 


## Calculation of rCSI Indicator & CH 

### Example data set

Here is the example data set:
[datarCSIEng](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/example_datasets/datarCSIEng.sav) 

### SPSS Syntax

```{r, eval = F}

GET  FILE='datarCSIEng.sav'.
DATASET NAME DataSet1 WINDOW=FRONT.

** caculate rCSI

compute rCSI = sum(rCSILessQlty,rCSIBorrow*2,rCSIMealSize,rCSIMealAdult*3,rCSIMealNb).
Variable labels rCSI "rCSI".

** each household should have a rCSI between 0 - 56

FREQUENCIES VARIABLES =  rCSI
/STATISTICS=MEAN MEDIAN MINIMUM MAXIMUM
/ORDER=ANALYSIS.

** Create rCSI Cadre Harmonise Categories

RECODE rCSI (0 thru 3=1) (4 thru 18=2) (19 thru 56=3) INTO rCSI_CH.
variable labels rCSI_CH "rCSI categories - Cadre Harmonise ".
value labels rCSI_CH
1 `Phase1`
2 `Phase2`
3 `Phase3`.

** Generate table of proportion of households in CH HHhS phases by Adm1 and Adm2 using weights

WEIGHT BY WeightHH.

CROSSTABS
  /TABLES=ADMIN2Name BY rCSI_CH BY ADMIN1Name
  /FORMAT=AVALUE TABLES
  /CELLS=ROW 
  /COUNT ROUND CELL.


```

Here is the SPSS syntax file: 

[RBDstandardized_SPSSsyntaxrCSI](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/syntax/RBDstandardized_SPSSsyntaxrCSI.sps) 


### R Syntax

```{r, eval = T, message = F}
library(haven)
library(labelled)
library(tidyverse)

#import dataset
datarCSIEng <- read_sav("example_datasets/datarCSIEng.sav")

#Calculate HHS 
datarCSIEng <- to_factor(datarCSIEng)

#calculate rCSI Score
datarCSIEng <- datarCSIEng %>% mutate(rCSI = rCSILessQlty  + (2 * rCSIBorrow) + rCSIMealSize + (3 * rCSIMealAdult) + rCSIMealNb)
var_label(datarCSIEng$rCSI) <- "rCSI"

#each household should have a rCSI between 0 - 56

summary(datarCSIEng$rCSI)

#Create rCSI Cadre Harmonise Categories
datarCSIEng <- datarCSIEng %>% mutate(rCSI_CH = case_when(
  rCSI <= 3 ~ "Phase1", 
  between(rCSI,4,18) ~ "Phase2",       
  rCSI >= 19 ~ "Phase3"))
var_label(datarCSIEng$rCSI_CH) <- "rCSI categories - Cadre Harmonise "

#Generate table of proportion of households in CH rCSI phases by Adm1 and Adm2 using weights
CH_rCSI_table_wide <- datarCSIEng %>% group_by(ADMIN1Name, ADMIN2Name) %>%
  drop_na(rCSI_CH) %>%
  count(rCSI_CH, wt = WeightHH) %>%
  mutate(perc = 100 * n / sum(n)) %>%
  ungroup() %>% select(-n) %>%
  spread(key = rCSI_CH, value = perc) %>% replace(., is.na(.), 0)  %>% mutate_if(is.numeric, round, 1)




```

Here is the R syntax file: 

[RBDstandardized_RsyntaxrCSI](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/syntax/RBDstandardized_RsyntaxrCSI.R) 


## Data Quality Monitoring & Visualisation of HHS Indicator

### Calculating & Visualizing CH finali phasing HHS Indicator for an Area 











### Monitoring 

1. Compare HHS CH phasing by admin1 and admin2 areas

```{r, eval = T, message = F}
library(plotly)

#Generate longtable of proportion of households in CH rCSI phases by Adm1 and Adm2 not using weights
CH_rCSI_table_long <- datarCSIEng %>% group_by(ADMIN1Name, ADMIN2Name) %>%
  drop_na(rCSI_CH) %>%
  count(rCSI_CH) 

#make graph
CH_rCSI_admin2_bargraph <- CH_rCSI_table_long %>% ggplot(aes(fill=rCSI_CH, y=n, x=ADMIN2Name)) +geom_bar(position="fill", stat="identity") +facet_wrap(. ~ ADMIN1Name, scales = "free") +scale_fill_manual(values=c("#c6ffc7","#ffe718","#e88400","#e02d00", "#5e0803")) +theme_minimal() +theme(axis.text.x = element_text(angle = 90)) +scale_y_continuous(labels = scales::percent)+ theme(axis.title.x = element_blank(), axis.title.y = element_blank()) 

#add title
CH_rCSI_admin2_bargraph <- CH_rCSI_admin2_bargraph +ggtitle("CH Phasing of rCSI Groups of Households by Admin2 Area") +theme(plot.title = element_text(hjust = 0.5))

ggplotly(CH_rCSI_admin2_bargraph)
```


2. Examine HHS CH phasing by admin2 areas and enumerators 

```{r, eval = T, message = F}
library(plotly)

#Generate longtable of proportion of households in CH rCSI phases by Adm1 and Adm2 and Operator - not using survey weights 
CH_rCSI_enumerators_tablelong <- datarCSIEng   %>% 
  group_by(ADMIN2Name, enum_cod) %>%
  count(rCSI_CH) 


#make graph
CH_rCSI_admin2enumerators_bargraph <- CH_rCSI_enumerators_tablelong %>% ggplot(aes(fill=rCSI_CH, y=n, x=enum_cod)) +geom_bar(position="fill", stat="identity") +facet_wrap(. ~ ADMIN2Name, scales = "free") +scale_fill_manual(values=c("#c6ffc7","#ffe718","#e88400","#e02d00", "#5e0803")) +theme_void()

#add title
CH_rCSI_admin2enumerators_bargraph  <- CH_rCSI_admin2enumerators_bargraph +ggtitle("CH Phasing of HHS of Households by Admin2 Area and Enumerator") +theme(plot.title = element_text(hjust = 0.5))

ggplotly(CH_rCSI_admin2enumerators_bargraph)

```
