# Food Consumption Score & Household Dietary Diversity

The modules below allow you to collect and calculate FCS, FCS-N and HDDS indicators.

## References / Resources

1.  FCS - https://documents.wfp.org/stellent/groups/public/documents/manual_guide_proced/wfp271449.pdf

2.  FCS_N - https://documents.wfp.org/stellent/groups/public/documents/manual_guide_proced/wfp277333.pdf

3.  HDDS - http://www.fao.org/fileadmin/user_upload/wa_workshop/docs/FAO-guidelines-dietary-diversity2011.pdf

## Standardized Questionnaire

### Paper Version of Questionnaire

Here is the standardized module in word format:

[RBDstandardized_questionnaireFCSHDDS](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/questionnaires/RBDstandardized_questionnaireFCSHDDS.docx) 

### Electronic Version of Questionnaire

Here is the standardized module in xlsform:

[RBDstandardized_questionnaireFCSHDDS](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/questionnaires/RBDstandardized_questionnaireFCSHDDS.xlsx)

## Calculation of FCS & HDDS Indicators  

### Example data set

Here is the example data set:
[dataFCSHDDSEng](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/example_datasets/dataFCSHDDSEng.sav) 

still need to create 

### SPSS Syntax

```{r, eval = F}

to be added

```

Here is the SPSS syntax file: 

not done yet

### R Syntax

```{r, eval = F}
library(haven)
library(labelled)
library(tidyverse)

#import dataset
dataFCSHDDSEng <- read_sav("example_datasets/dataFCSHDDSEng.sav")
#convert to labels
dataFCSHDDSEng <- to_factor(dataFCSHDDSEng)

#calculate FCS
dataFCSHDDSEng <- dataFCSHDDSEng %>% mutate(FCS = (2 * FCSStap) + (3 * FCSPulse)+ (4*FCSPr) +FCSVeg  +FCSFruit +(4*FCSDairy) + (0.5*FCSFat) + (0.5*FCSSugar))
#create FCG groups based on 21/25 or 28/42 thresholds
dataFCSHDDSEng <- dataFCSHDDSEng %>% mutate(
  FCSCat21 = case_when(
  FCS <= 21 ~ "Poor", between(FCS, 21.5, 35) ~ "Borderline", FCS > 35 ~ "Acceptable"),
  FCSCat28 = case_when(
  FCS <= 28 ~ "Poor", between(FCS, 28.5, 42) ~ "Borderline", FCS > 42 ~ "Acceptable"))
var_label(dataFCSHDDSEng$FCSCat21) <- "Food Consumption Group 21/35 thresholds"
var_label(dataFCSHDDSEng$FCSCat28) <-  "Food Consumption Group 28/42 thresholds"



#calculate HDDS first by creating the 12 groups based on the 16 questions 
dataFCSHDDSEng <- dataFCSHDDSEng %>% mutate(
    HDDSStapCer = case_when(HDDSStapCer == "Yes" ~ 1, TRUE ~ 0),
    HDDSStapRoot = case_when(HDDSStapRoot  == "Yes" ~ 1, TRUE ~ 0),
    HDDSVeg = case_when(HDDSVegOrg  == "Yes" | HDDSVegGre == "Yes" | HDDSVegOth == "Yes" ~ 1, TRUE ~ 0),
    HDDSFruit = case_when(HDDSFruitOrg == "Yes" | HDDSFruitOth == "Yes" ~ 1, TRUE ~ 0),
    HDDSPrMeat = case_when(HDDSPrMeatF == "Yes" | HDDSPrMeatO == "Yes" ~ 1, TRUE ~ 0),
    HDDSPrEgg = case_when(HDDSPrEgg  == "Yes" ~ 1, TRUE ~ 0),
    HDDSPrFish = case_when(HDDSPrFish == "Yes" ~ 1, TRUE ~ 0),
    HDDSPulse = case_when(HDDSPulse == "Yes" ~ 1, TRUE ~ 0),
    HDDSDairy = case_when(HDDSDairy == "Yes" ~ 1, TRUE ~ 0),
    HDDSFat = case_when(HDDSFat == "Yes" ~ 1, TRUE ~ 0),
    HDDSSugar = case_when(HDDSSugar == "Yes" ~ 1, TRUE ~ 0),
    HDDSCond = case_when(HDDSCond == "Yes"~ 1, TRUE ~ 0))

#Calculate HDDS and Cadre Harmonise Phases
dataFCSHDDSEng <- dataFCSHDDSEng %>% mutate(HDDS = HDDSStapCer +HDDSStapRoot +HDDSVeg +HDDSFruit +HDDSPrMeat +HDDSPrEgg +HDDSPrFish +HDDSPulse +HDDSDairy +HDDSFat +HDDSSugar +HDDSCond)
var_label(dataFCSHDDSEng$HDDS) <- "Hosehold Dietary Diversity Score"

#Calucate Cadre Harmonise HDDS phasing categories 
dataFCSHDDSEng <- dataFCSHDDSEng %>%  mutate(HDDS_CH = case_when(
    HDDS >= 5 ~ "Phase1", 
    HDDS == 4 ~ "Phase2",       
    HDDS == 3 ~ "Phase3",
    HDDS == 2 ~ "Phase4",
    HDDS < 2 ~ "Phase5"))
var_label(dataFCSHDDSEng$HDDS_CH) <- "Hosehold Dietary Diversity Score - CH phasing"

#Generate table of proportion of households in FCG by Adm1 and Adm2 using weights 
#Food Consumption Group 21/35 cutoff 
FCSCat_table_wide <- dataFCSHDDSEng %>% 
  drop_na(FCSCat21) %>%
  group_by(ADMIN1Name, ADMIN2Name, ADMIN2Code) %>%
  count(FCSCat21, wt = WeightHH) %>%
  mutate(perc = 100 * n / sum(n)) %>%
  ungroup() %>% select(-n) %>%
  spread(key = FCSCat21, value = perc) %>% replace(., is.na(.), 0) %>% mutate_if(is.numeric, round, 1)

#Calculate phasing of FCG indicator for area (applying CH 20% rules)
FCSCat_table_wide <- FCSCat_table_wide %>%
  mutate(PoorBorderline = Poor + Borderline, FCG_finalphase = case_when(
  Poor < 5 ~ 1,  #if less than 5% are in the poor food group then phase 1
  Poor >= 20 ~ 4, #if 20% or more are in the poor food group then phase 4
  between(Poor,5,10) ~ 2, #if % of people are between 5 and 10%  then phase2
  between(Poor,10,20) & PoorBorderline < 30 ~ 2, #if % of people in poor food group are between 20 and 20% and the % of people who are in poor and borderline is less than 30 % then phase2
  between(Poor,10,20) & PoorBorderline >= 30 ~ 3)) %>% #if % of people in poor food group are between 20 and 20% and the % of people who are in poor and borderline is less than 30 % then phase2
  select(ADMIN1Name, ADMIN2Name, ADMIN2Code, FCG_Poor = Poor, FCG_Borderline = Borderline, FCG_Acceptable = Acceptable, FCG_finalphase) #select only relevant variables and order in proper sequence

#Generate table of proportion of households by HDDS by Adm1 and Adm2 using weights
CH_HDDS_table_wide <- dataFCSHDDSEng %>% 
  drop_na(HDDS_CH) %>%
  group_by(ADMIN1Name, ADMIN2Name, ADMIN2Code) %>%
  count(HDDS_CH, wt = WeightHH) %>%
  mutate(perc = 100 * n / sum(n)) %>%
  ungroup() %>% select(-n) %>%
  spread(key = HDDS_CH, value = perc) %>% replace(., is.na(.), 0) %>% mutate_if(is.numeric, round, 1)

#Calculate phasing of HDDS indicator for area (applying CH 20% rules)


#Apply the 20% rule (if it is 20% in that phase or the sum of higher phases equals 20%) 
CH_HDDS_table_wide <- CH_HDDS_table_wide %>% 
    mutate(
    phase2345 = `Phase2` + `Phase3` + `Phase4` + `Phase5`, 
    phase345 = `Phase3` + `Phase4` + `Phase5`,
    phase45 = `Phase4` + `Phase5`, 
    HDDS_finalphase = case_when(
    `Phase5` >= 20 ~ 5, 
    `Phase4` >= 20 | phase45 >= 20 ~ 4, 
    `Phase3` >= 20 | phase345 >= 20 ~ 3, 
    `Phase2` >= 20 | phase2345 >= 20 ~ 2,
     TRUE ~ 1)) %>% 
  select(ADMIN1Name, ADMIN2Name, ADMIN2Code, HDDS_Phase1 = Phase1, HDDS_Phase2 = Phase2, HDDS_Phase3 = Phase3, HDDS_Phase4 = Phase4, HDDS_Phase5 = Phase5, HDDS_finalphase) 
```


Here is the R syntax file: 

[RBDstandardized_RsyntaxFCSHDDS](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/syntax/RBDstandardized_RsyntaxFCSHDDS.R) 


## Data Quality Monitoring & Visualisation of FCS Indicator

### Mapping FCS final phasing

```{r, eval = T, message = F}
library(sf)

#import shapefile of Nigeria
NGA_shp2 <- read_sf("example_datasets/NGA_shpfile/nga_admbnda_adm2_osgof_20190417.shp") %>% filter(ADM1_EN %in% c("Adamawa","Borno","Yobe"))
NGA_shp1 <- read_sf("example_datasets/NGA_shpfile/nga_admbnda_adm1_osgof_20190417.shp") %>% filter(ADM1_REF %in% c("Adamawa","Borno","Yobe"))
#rename 
NGA_shp2 <- NGA_shp2 %>% rename("ADMIN2Code" = "ADM2_PCODE")
#join shapefile to table of HHS phases
CH_FCG_shp <- left_join(NGA_shp2, FCSCat_table_wide, by = "ADMIN2Code")

#CH color codes
CH_colors = c("1" = "#c6ffc7", "2" = "#ffe718", "3" = "#e88400", "4" = "#e02d00", "5" = "#5e0803")

mapFCG <- CH_FCG_shp %>% ggplot()  +geom_sf(aes(fill = as.factor(FCG_finalphase))) +theme_void() +coord_sf(datum=NA) +labs(fill="FCG phasing") +scale_fill_manual(values = CH_colors) +ggtitle(label= "phasing of FCG") +geom_sf(data=NGA_shp1,  lwd=1.5, fill=NA) 

mapFCG
```






