# Food Consumption Score & Household Dietary Diversity

The modules below allow you to collect and calculate FCS, FCS-N and HDDS indicators.

## Standardized Questionnaire

### Paper Version of Questionnaire

Here is the standardized module in word format:

[RBDstandardized_questionnaireFCSHDDS](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/questionnaires/RBDstandardized_questionnaireFCSHDDS.docx) 

### Electronic Version of Questionnaire

Here is the standardized module in xlsform:

[RBDstandardized_questionnaireFCSHDDS](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/questionnaires/RBDstandardized_questionnaireFCSHDDS.xlsx)

## Calculation of FCS & HDDS Indicators  

### Example data set

Here is the example data set:
[dataFCSHDDSEng](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/example_datasets/dataFCSHDDSEng.sav) 

still need to create 

### SPSS Syntax

```{r, eval = F}

to be added

```

Here is the SPSS syntax file: 

not done yet

### R Syntax

```{r, eval = F}
library(haven)
library(labelled)
library(tidyverse)

#import dataset
dataFCSHDDSEng <- read_sav("example_datasets/dataFCSHDDSEng.sav")
#convert to labels
dataFCSHDDSEng <- to_factor(dataFCSHDDSEng)

#calculate FCS
dataFCSHDDSEng <- dataFCSHDDSEng %>% mutate(FCS = (2 * FCSStap) + (3 * FCSPulse)+ (4*FCSPr) +FCSVeg  +FCSFruit +(4*FCSDairy) + (0.5*FCSFat) + (0.5*FCSSugar))
#create FCG groups based on 21/25 or 28/42 thresholds
dataFCSHDDSEng <- dataFCSHDDSEng %>% mutate(
  FCSCat21 = case_when(
  FCS <= 21 ~ "Poor", between(FCS, 21.5, 35) ~ "Borderline", FCS > 35 ~ "Acceptable"),
  FCSCat28 = case_when(
  FCS <= 28 ~ "Poor", between(FCS, 28.5, 42) ~ "Borderline", FCS > 42 ~ "Acceptable"))
var_label(dataFCSHDDSEng$FCSCat21) <- "Food Consumption Group 21/35 thresholds"
var_label(dataFCSHDDSEng$FCSCat28) <-  "Food Consumption Group 28/42 thresholds"



#calculate HDDS first by creating the 12 groups based on the 16 questions 
dataFCSHDDSEng <- dataFCSHDDSEng %>% mutate(
    HDDSStapCer = case_when(HDDSStapCer == "Yes" ~ 1, TRUE ~ 0),
    HDDSStapRoot = case_when(HDDSStapRoot  == "Yes" ~ 1, TRUE ~ 0),
    HDDSVeg = case_when(HDDSVegOrg  == "Yes" | HDDSVegGre == "Yes" | HDDSVegOth == "Yes" ~ 1, TRUE ~ 0),
    HDDSFruit = case_when(HDDSFruitOrg == "Yes" | HDDSFruitOth == "Yes" ~ 1, TRUE ~ 0),
    HDDSPrMeat = case_when(HDDSPrMeatF == "Yes" | HDDSPrMeatO == "Yes" ~ 1, TRUE ~ 0),
    HDDSPrEgg = case_when(HDDSPrEgg  == "Yes" ~ 1, TRUE ~ 0),
    HDDSPrFish = case_when(HDDSPrFish == "Yes" ~ 1, TRUE ~ 0),
    HDDSPulse = case_when(HDDSPulse == "Yes" ~ 1, TRUE ~ 0),
    HDDSDairy = case_when(HDDSDairy == "Yes" ~ 1, TRUE ~ 0),
    HDDSFat = case_when(HDDSFat == "Yes" ~ 1, TRUE ~ 0),
    HDDSSugar = case_when(HDDSSugar == "Yes" ~ 1, TRUE ~ 0),
    HDDSCond = case_when(HDDSCond == "Yes"~ 1, TRUE ~ 0))

#Calculate HDDS and Cadre Harmonise Phases
dataFCSHDDSEng <- dataFCSHDDSEng %>% mutate(HDDS = HDDSStapCer +HDDSStapRoot +HDDSVeg +HDDSFruit +HDDSPrMeat +HDDSPrEgg +HDDSPrFish +HDDSPulse +HDDSDairy +HDDSFat +HDDSSugar +HDDSCond)
var_label(dataFCSHDDSEng$HDDS) <- "Hosehold Dietary Diversity Score"

#Calucate Cadre Harmonise HDDS phasing categories 
dataFCSHDDSEng <- dataFCSHDDSEng %>%  mutate(HDDS_CH = case_when(
    HDDS >= 5 ~ "Phase1", 
    HDDS == 4 ~ "Phase2",       
    HDDS == 3 ~ "Phase3",
    HDDS == 2 ~ "Phase4",
    HDDS < 2 ~ "Phase5"))
var_label(dataFCSHDDSEng$HDDS_CH) <- "Hosehold Dietary Diversity Score - CH phasing"

#Generate table of proportion of households in FCG by Adm1 and Adm2 using weights 
#Food Consumption Group 21/35 cutoff 
FCSCat_table_wide <- dataFCSHDDSEng %>% 
  drop_na(FCSCat21) %>%
  group_by(ADMIN1Name, ADMIN2Name) %>%
  count(FCSCat21, wt = WeightHH) %>%
  mutate(perc = 100 * n / sum(n)) %>%
  ungroup() %>% select(-n) %>%
  spread(key = FCSCat21, value = perc) %>% replace(., is.na(.), 0) %>% mutate_if(is.numeric, round, 1)

#Generate table of proportion of households by HDDS by Adm1 and Adm2 using weights
#Food Consumption Group 21/35 cutoff 
CH_HDDS_table_wide <- dataFCSHDDSEng %>% 
  drop_na(HDDS_CH) %>%
  group_by(ADMIN1Name, ADMIN2Name) %>%
  count(HDDS_CH, wt = WeightHH) %>%
  mutate(perc = 100 * n / sum(n)) %>%
  ungroup() %>% select(-n) %>%
  spread(key = HDDS_CH, value = perc) %>% replace(., is.na(.), 0) %>% mutate_if(is.numeric, round, 1)
```


Here is the R syntax file: 

[RBDstandardized_RsyntaxFCSHDDS](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/syntax/RBDstandardized_RsyntaxFCSHDDS.R) 


## References / Resources

1.  FCS - https://documents.wfp.org/stellent/groups/public/documents/manual_guide_proced/wfp271449.pdf

2.  FCS_N - https://documents.wfp.org/stellent/groups/public/documents/manual_guide_proced/wfp277333.pdf

3.  HDDS - http://www.fao.org/fileadmin/user_upload/wa_workshop/docs/FAO-guidelines-dietary-diversity2011.pdf



