# Livelihood Coping Strategies

## Standardized Module - Light - Coping for Food

Variable Name | Question Label | Answer Choices
---- | ------------- |------- 
LhCSIStress1_lt | During the past 30 days, did anyone in your household have to: *Sell household assets/goods  radio, furniture, refrigerator, television, jewelry etc..)* due to a lack of food or a lack of money to buy food? | 1) No, because I did not face a shortage of food 2) No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it 3) Yes 4) Not applicable
LhCSIStress2_lt | During the past 30 days, did anyone in your household have to: *sell more animals (non-productive) than usual* due to a lack of food or a lack of money to buy food? | 1) No, because I did not face a shortage of food 2) No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it 3) Yes 4) Not applicable
LhCSIStress3_lt | During the past 30 days, did anyone in your household have to: *spend savings* due to a lack of food or a lack of money to buy food? | 1) No, because I did not face a shortage of food 2) No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it 3) Yes 4) Not applicable
LhCSIStress4_lt | During the past 30 days, did anyone in your household have to: *Borrow money / food  from a formal lender / bank* due to a lack of food or a lack of money to buy food? | 1) No, because I did not face a shortage of food 2) No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it 3) Yes 4) Not applicable
LhCSICrisis1_lt | During the past 30 days, did anyone in your household have to: *Reduce non-food expenses on health (including drugs) and education * due to a lack of food or a lack of money to buy food? | 1) No, because I did not face a shortage of food 2) No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it 3) Yes 4) Not applicable
LhCSICrisis2_lt | During the past 30 days, did anyone in your household have to: *sell productive assets or means of transport (sewing machine, wheelbarrow, bicycle, car, etc..)* due to a lack of food or a lack of money to buy food? | 1) No, because I did not face a shortage of food 2) No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it 3) Yes 4) Not applicable
LhCSICrisis3_lt | During the past 30 days, did anyone in your household have to: *withdraw children from school* due to a lack of food or a lack of money to buy food? | 1) No, because I did not face a shortage of food 2) No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it 3) Yes 4) Not applicable
LhCSIEmergency1_lt | During the past 30 days, did anyone in your household have to: *sell house or land* due to a lack of food or a lack of money to buy food? | 1) No, because I did not face a shortage of food 2) No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it 3) Yes 4) Not applicable
LhCSIEmergency2_lt | During the past 30 days, did anyone in your household have to: *sell last female animals* due to a lack of food or a lack of money to buy food? | 1) No, because I did not face a shortage of food 2) No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it 3) Yes 4) Not applicable
LhCSIEmergency3_lt | During the past 30 days, did anyone in your household have to: *beg * due to a lack of food or a lack of money to buy food? | 1) No, because I did not face a shortage of food 2) No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it 3) Yes 4) Not applicable

### Standardized Module - data collection form

Here is the standardized module in xlsform: [RBDstandardized_questionnaireLHCS](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/questionnaires/RBDstandardized_questionnaireLHCS.xlsx) 

### Analysis 

#### Example data set
[dataLHCSLightEng](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/example_datasets/dataLHCSLightEng.sav) 

Here is the example data set:

#### SPSS Syntax

```{r, eval = F}
```

Here is the SPSS syntax file: 

still need to do


#### R Syntax

```{r, eval = F}
library(haven)
library(labelled)
library(tidyverse)

#import dataset
dataLHCSLightEng <- read_sav("dataLHCSLightEng.sav")

#Calculate HHS 
dataLHCSLightEng <- to_factor(dataLHCSLightEng)

#create a variable to specify if the household used any of the strategies by severity 
#stress
dataLHCSLightEng <- dataLHCSLightEng %>% mutate(stress_coping = case_when(
LhCSIStress1 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedstress",
LhCSIStress2 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedstress",
LhCSIStress3 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedstress",
LhCSIStress4 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedstress",
TRUE ~ "notusedstress"))
#Crisis
dataLHCSLightEng <- dataLHCSLightEng %>% mutate(crisis_coping = case_when(
LhCSICrisis1 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedcrisis",
LhCSICrisis2 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedcrisis",
LhCSICrisis3 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedcrisis",
TRUE ~ "notusedcrisis"))
#Emergency
dataLHCSLightEng <- dataLHCSLightEng %>% mutate(emergency_coping = case_when(
LhCSIEmergency1 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedemergency",
LhCSIEmergency2 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedemergency",
LhCSIEmergency3 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedemergency",
TRUE ~ "noutusedemergency"))

#calculate Max_coping_behaviour
dataLHCSLightEng <- dataLHCSLightEng %>% mutate(LhCSICat = case_when(
  emergency_coping == "usedemergency" ~ "EmergencyStrategies",
  crisis_coping == "usedcrisis" ~ "CrisisStrategies",
  stress_coping == "usedstress" ~ "StressStrategies",
  TRUE ~ "NoStrategies"))

#Generate table of proportion of households in CH LhCHS phases by Adm1 and Adm2 using weights
#Livelihood Coping Strategies 
LhHCSCat_table_wide <- dataLHCSLightEng %>% 
  drop_na(LhCSICat) %>%
  group_by(ADMIN1Name, ADMIN2Name) %>%
  count(LhCSICat, wt = WeightHH) %>%
  mutate(perc = 100 * n / sum(n)) %>%
  ungroup() %>% select(-n) %>%
  spread(key = LhCSICat, value = perc) %>% replace(., is.na(.), 0) %>% mutate_if(is.numeric, round, 1)
```


Here is the R syntax file: 
[RBDstandardized_RsyntaxLHCSIlight](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/syntax/RBDstandardized_RsyntaxLHCSIlight.R) 


## Standardized Module - Light - Essential Needs Assessment (ENA)

Variable Name | Question Label | Answer Choices
---- | ------------- |------- 
LhCSIStress1_EN | During the past 30 days, did anyone in your household have to: Sell household assets/goods (radio, furniture, refrigerator, television, jewelry etc..) because there was not enough resources (food, cash, else) to meet essential needs (e.g. adequate shelter, education services, health services, food, etc)? | 1) No, because I did not face a shortage of food 2) No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it 3) Yes 4) Not applicable
LhCSIStress1_EN_why | What was the MAIN REASON you or other members in your household adopted these coping strategies to access essential needs? | 1)To access food 2)To access education services/commodities (e.g. uniforms, books) 3)To access health services/medicines 4)To buy and repair cloths or shoes 5)To access adequate shelter 6)To access water/sanitation facilities 7)To access essential dwelling services (electricity, energy, waste disposal…) 8)To pay for existing debts 9)Other  

### Standardized Module - data collection form

Here is the standardized module in xlsform: [RBDstandardized_questionnaireLHCS](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/questionnaires/RBDstandardized_questionnaireLHCS.xlsx) 

### Analysis 

#### Example data set
[dataLHCSLightEng](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/example_datasets/dataLHCSLightEng.sav) 

Here is the example data set:

#### SPSS Syntax

```{r, eval = F}
```

Here is the SPSS syntax file: 

still need to do


#### R Syntax

```{r, eval = F}
library(haven)
library(labelled)
library(tidyverse)

#import dataset
dataLHCSLightEng <- read_sav("dataLHCSLightEng.sav")

#Calculate HHS 
dataLHCSLightEng <- to_factor(dataLHCSLightEng)

#create a variable to specify if the household used any of the strategies by severity 
#stress
dataLHCSLightEng <- dataLHCSLightEng %>% mutate(stress_coping = case_when(
LhCSIStress1 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedstress",
LhCSIStress2 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedstress",
LhCSIStress3 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedstress",
LhCSIStress4 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedstress",
TRUE ~ "notusedstress"))
#Crisis
dataLHCSLightEng <- dataLHCSLightEng %>% mutate(crisis_coping = case_when(
LhCSICrisis1 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedcrisis",
LhCSICrisis2 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedcrisis",
LhCSICrisis3 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedcrisis",
TRUE ~ "notusedcrisis"))
#Emergency
dataLHCSLightEng <- dataLHCSLightEng %>% mutate(emergency_coping = case_when(
LhCSIEmergency1 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedemergency",
LhCSIEmergency2 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedemergency",
LhCSIEmergency3 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedemergency",
TRUE ~ "noutusedemergency"))

#calculate Max_coping_behaviour
dataLHCSLightEng <- dataLHCSLightEng %>% mutate(LhCSICat = case_when(
  emergency_coping == "usedemergency" ~ "EmergencyStrategies",
  crisis_coping == "usedcrisis" ~ "CrisisStrategies",
  stress_coping == "usedstress" ~ "StressStrategies",
  TRUE ~ "NoStrategies"))

#Generate table of proportion of households in CH LhCHS phases by Adm1 and Adm2 using weights
#Livelihood Coping Strategies 
LhHCSCat_table_wide <- dataLHCSLightEng %>% 
  drop_na(LhCSICat) %>%
  group_by(ADMIN1Name, ADMIN2Name) %>%
  count(LhCSICat, wt = WeightHH) %>%
  mutate(perc = 100 * n / sum(n)) %>%
  ungroup() %>% select(-n) %>%
  spread(key = LhCSICat, value = perc) %>% replace(., is.na(.), 0) %>% mutate_if(is.numeric, round, 1)
```


Here is the R syntax file: 
[RBDstandardized_RsyntaxLHCSIlight](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/syntax/RBDstandardized_RsyntaxLHCSIlight.R) 

