--- 
title: "A Guide to Collection & Calculation of Cadre Harmonise Food Security Indicators - beta version"
author: "Research, Assesment and Monitoring (RAM) Team
         Regional Bureau Dakar
         World Food Programme
         rbd.vam@wfp.org"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
---

# Purpose & How to use this Guide

This guide proposes a uniform way of collecting and analyzing key food security indicators for the Cadre Harmonise process.

Chapters are listed by indicator and include:

1. standardized data collection modules and ODK compliant xlsforms
2. sample data set 
3. standardized syntax in SPSS and R
4. references / official guidelines / source materials for the indicators



<!--chapter:end:index.Rmd-->

# Household Hunger Scale {#HHS}

## Standardized Questionnaire

Variable Name | Question Label | Answer Choices
---- | ------------- |---- 
HHhSNoFood_YN | In the past [4 weeks/30 days], was there ever no food to eat of any kind in your house because of lack of resources to get food? | 1) Yes 0) No
HHhSNoFood_FR | How often did this happen in the past [4 weeks/30 days]? | 1) Rarely (1–2 times) 2) Sometimes (3–10 times)  3) Often (more than 10 times)
HHhSBedHung_YN | In the past [4 weeks/30 days], did you or any household member go to sleep at night hungry because there was not enough food? | 1) Yes 0) No
HHhSBedHung_FR  | How often did this happen in the past [4 weeks/30 days]? | 1) Rarely (1–2 times) 2) Sometimes (3–10 times)  3) Often (more than 10 times)
HHhSNotEat_YN | In the past [4 weeks/30 days], did you or any household member go to sleep at night hungry because there was not enough food? | 1) Yes 0) No
HHhSNotEat_FR | How often did this happen in the past [4 weeks/30 days]? | 1) Rarely (1–2 times) 2) Sometimes (3–10 times)  3) Often (more than 10 times)

Here is the standardized module in xlsform:

[RBDstandardized_questionnaireHHS](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/questionnaires/RBDstandardized_questionnaireHHS.xlsx) 

## Calculation of HHS Indicator & CH 

### Example data set

Here is the example data set:
[dataHHSEng](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/example_datasets/dataHHSEng.sav) 

### SPSS Syntax

```{r, eval = F}

to be added

```

Here is the SPSS syntax file: 

not done yet

### R Syntax

```{r, eval = F}
library(haven)
library(labelled)
library(tidyverse)

#import dataset
dataHHSEng <- read_sav("dataHHSEng.sav")

#Calculate HHS 
dataHHSEng <- to_factor(dataHHSEng)

#Recode HHS questions into new variables with score 
dataHHSEng <- dataHHSEng %>% mutate(HHhSNoFood_FR_r = case_when(
                                      HHhSNoFood_FR == "Rarely (1–2 times)" ~ 1,
                                      HHhSNoFood_FR == "Sometimes (3–10 times)" ~ 1,
                                      HHhSNoFood_FR == "Often (more than 10 times)" ~ 2,
                                      TRUE ~ 0),
                                    HHhSBedHung_FR_r = case_when(
                                      HHhSBedHung_FR == "Rarely (1–2 times)" ~ 1,
                                      HHhSBedHung_FR == "Sometimes (3–10 times)" ~ 1,
                                      HHhSBedHung_FR == "Often (more than 10 times)" ~ 2,
                                      TRUE ~ 0),
                                    HHhSNotEat_FR_r = case_when(
                                      HHhSNotEat_FR == "Rarely (1–2 times)" ~ 1,
                                      HHhSNotEat_FR == "Sometimes (3–10 times)" ~ 1,
                                      HHhSNotEat_FR == "Often (more than 10 times)" ~ 2,
                                      TRUE ~ 0))
#Calculate HHS
dataHHSEng <- dataHHSEng %>% mutate(HHhS = HHhSNoFood_FR_r + HHhSBedHung_FR_r + HHhSNotEat_FR_r)

#Convert HH Scores to CH phases
dataHHSEng <- dataHHSEng %>% mutate(CH_HHS = case_when(
    HHhS == 0 ~ "Phase1",
    HHhS == 1 ~ "Phase2",
    HHhS %in% c(2,3,4) ~ "Phase3",
    HHhS == 4 ~ "Phase4",  
    HHhS >= 5 ~ "Phase5"))

#Generate table of proportion of households in CH HHS phases by Adm1 and Adm2 using weights
CH_HHS_table_wide <- dataHHSEng %>% group_by(ADMIN1Name, ADMIN2Name) %>%
  drop_na(CH_HHS) %>%
  count(CH_HHS, wt = hh_weight) %>%
  mutate(perc = 100 * n / sum(n)) %>%
  ungroup() %>% select(-n) %>%
  spread(key = CH_HHS, value = perc) %>% replace(., is.na(.), 0)  %>% mutate_if(is.numeric, round, 1)
```


Here is the R syntax file: 

[RBDstandardized_RsyntaxHHS](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/syntax/RBDstandardized_RsyntaxHHS.R) 


## References / Resources



<!--chapter:end:01-HHS.Rmd-->

# reduced Coping Strategy Index {#rCSI}

## Standardized Questionnaire

Variable Name | Question Label 
---- | ------------- 
rCSILessQlty | In the past 7 days, how many days has your household had to: *Rely on less preferred and less expensive food* because you did not have enough food or money to buy food?
rCSIBorrow | In the past 7 days, how many days has your household had to: *Borrow food or rely on help from a relative or friend* because you did not have enough food or money to buy food?
rCSIMealSize | In the past 7 days, how many days has your household had to: *Limit portion size of meals* because you did not have enough food or money to buy food?
rCSIMealAdult  | In the past 7 days, how many days has your household had to: *Restrict consumption by adults in order for small children to eat* because you did not have enough food or money to buy food?
rCSIMealNb | In the past 7 days, how many days has your household had to: *Reduce the number of meals eaten per day* because you did not have enough food or money to buy food?

Here is the standardized module in xlsform:

[RBDstandardized_questionnaireCSI](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/questionnaires/RBDstandardized_questionnairerCSI.xlsx) 


## Calculation of rCSI Indicator & CH 

### Example data set

Here is the example data set:
[datarCSIEng](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/example_datasets/dataHHSEng.sav) 

### SPSS Syntax

```{r, eval = F}

to be added

```

Here is the SPSS syntax file: 

not done yet

### R Syntax

```{r, eval = F}
library(haven)
library(labelled)
library(tidyverse)

#import dataset
datarCSIEng <- read_sav("datarCSIEng.sav")

#Calculate HHS 
datarCSIEng <- to_factor(datarCSIEng)

#calculate rCSI Score
datarCSIEng <- datarCSIEng %>% mutate(rCSI = rCSILessQlty  + (2 * rCSIBorrow) + rCSIMealSize + (3 * rCSIMealAdult) + rCSIMealNb)

#convert rCSI Score into thresholds 
datarCSIEng <- datarCSIEng %>% mutate(CH_rCSI = case_when(
  rCSI <= 3 ~ "Phase1", 
  between(rCSI,4,18) ~ "Phase2",       
  rCSI >= 19 ~ "Phase3"))

#Generate table of proportion of households in CH rCSI phases by Adm1 and Adm2 using weights
CH_rCSI_table_wide <- datarCSIEng %>% group_by(ADMIN1Name, ADMIN2Name) %>%
  drop_na(CH_rCSI) %>%
  count(CH_rCSI, wt = hh_weight) %>%
  mutate(perc = 100 * n / sum(n)) %>%
  ungroup() %>% select(-n) %>%
  spread(key = CH_rCSI, value = perc) %>% replace(., is.na(.), 0)  %>% mutate_if(is.numeric, round, 1)
```



Here is the R syntax file: 

[RBDstandardized_RsyntaxrCSI](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/syntax/RBDstandardized_RsyntaxrCSI.R) 


## References / Resources


<!--chapter:end:02-rCSI.Rmd-->

# Food Consumption Score & Household Dietary Diversity

## Standardized Questionnaire

Variable Name | Question Label 
---- | ------------- 
FCSStap | Over the last 7 days, how many days did members of your household eat : Cereals, grains, roots and tubers, such as: (Rice, pasta, bread, sorghum, millet, maize, fonio, potato, yam, cassava, white sweet potato) HDDSStapCer | Yesterday, during the day and night, did members in your household eat Cereals, such as: (cereals, grains,rice, pasta, bread, sorghum, millet, maize, fonio) 
FCSStapCerSRf | Over the last 7 days, what was the main source of Cereals, such as:  (cereals, grains, rice, pasta, bread, sorghum, millet, maize, fonio) 
HDDSStapRoot | Yesterday, during the day and night, did members in your household eat Roots and Tubers, such as: (potato, yam, cassava, white sweet potato)
FCSStapRootSRf | Over the last 7 days, what was the main source of Roots and Tubers, such as: (potato, yam, cassava, white sweet potato) 
FCSPulse | How many days over the last 7 days, did members of your household eat Pulses, such as: (beans, cowpeas, peanuts, lentils, nut, soy, pigeon pea and / or other nuts) 
HDDSPulse | Yesterday, during the day and night, did members in your household eat Pulses, such as: (beans, cowpeas, peanuts, lentils, nut, soy, pigeon pea and / or other nuts) 
FCSPulseSRf | Over the last 7 days, What was the main source of Pulses, such as: (beans, cowpeas, peanuts, lentils, nut, soy, pigeon pea and / or other nuts) 
FCSDairy | Over the last 7 days, how many days did members of your household eat/drink Milk and dairy products, such as:(fresh milk / sour, yogurt, cheese, other dairy products) excluding margarine / butter or small amounts of milk for tea / coffee ?
HDDSDairy | Yesterday, during the day and night, did members in your household eat/drink Milk and dairy products, such as: (fresh milk / sour, yogurt, cheese, other dairy products) excluding margarine / butter or small amounts of milk for tea / coffee ?
FCSDairySRf | Over the last 7 days, what was the main source of Milk and dairy products, such as: (fresh milk / sour, yogurt, cheese, other dairy products)  replace with locally relevant examples excluding margarine / butter or small amounts of milk for tea / coffee ?
FCSPr | Over the last 7 days, how many days did members of your household eat: Meat, fish, eggs, such as: (goat, beef, chicken, pork, blood, fish, including canned tuna, escargot, and / or other seafood, eggs) consumed in large quantities and not as a condiment ?
FCSPrMeatF | Over the last 7 days, how many days did members of your household eat: Flesh/red meat, such as: (beef, pork, lamb, goat, rabbit, chicken, duck, other birds, insects) consumed in large quantities and not as a condiment ?
HDDSPrMeatF | Yesterday, during the day and night, did members in your household eat: Flesh/red meat , such as: (beef, pork, lamb, goat, rabbit, chicken, duck, other birds, insects) consumed in large quantities and not as a condiment ?
FCSPrMeatFSRf | Over the last 7 days,What was the main source of Flesh/red meat , such as: (beef, pork, lamb, goat, rabbit, chicken, duck, other birds, insects) consumed in large quantities and not as a condiment ?
FCSPrMeatO | Over the last 7 days, how many days did members of your household eat: Organ meat, such as: (liver, kidney, heart and / or other organ meats) consumed in large quantities and not as a condiment ?
HDDSPrMeatO | Yesterday, during the day and night, did members in your household eat: Organ meat, such as: (liver, kidney, heart and / or other organ meats) consumed in large quantities and not as a condiment ?
FCSPrMeatOSRf | Over the last 7 days,What was the main source of Organ meat, such as: (liver, kidney, heart and / or other organ meats) consumed in large quantities and not as a condiment ?
FCSPrFish | Over the last 7 days, how many days did members of your household eat: Fish, such as: (fish, including canned tuna, escargot, and / or other seafood) consumed in large quantities and not as a condiment ?
HDDSPrFish | Yesterday, during the day and night, did members in your household eat: Fish, such as: (fish, including canned tuna, escargot, and / or other seafood) consumed in large quantities and not as a condiment ?
FCSPrFishSRf | Over the last 7 days,What was the main source of Fish, such as: (fish, including canned tuna, escargot, and / or other seafood) consumed in large quantities and not as a condiment ?
FCSPrEgg | How many days over the last 7 days, did members of your household eat : Eggs consumed in large quantities and not as a condiment ?
HDDSPrEgg | Yesterday, during the day and night, did members in your household eat: Eggs consumed in large quantities and not as a condiment ?
FCSPrEggSRf | Over the last 7 days,What was the main source of Eggs consumed in large quantities and not as a condiment ?
FCSVeg | Over the last 7 days, how many days did members of your household eat: Vegetables such as (spinach, onion, tomatoes, carrots, peppers, green beans, lettuce, etc.) ?
FCSVegOrg | Over the last 7 days, how many days did members of your household eat: Orange vegetables , such as: (carrot, red pepper, pumpkin, orange sweet potatoes, etc.) ?
HDDSVegOrg | Yesterday, during the day and night, did members in your household eat: Orange vegetables , such as: (carrot, red pepper, pumpkin, orange sweet potatoes, etc.) ?
FCSVegOrgSRf | Over the last 7 days,What was the main source of: Orange vegetables , such as: (carrot, red pepper, pumpkin, orange sweet potatoes, etc.) ?
FCSVegGre | Over the last 7 days, how many days did members of your household eat: Green leafy vegetables, such as: (spinach, broccoli, amaranth and / or other dark green leaves, cassava leaves, etc.) ?
HDDSVegGre | esterday, during the day and night, did members in your household eat: Green leafy vegetables, such as: (spinach, broccoli, amaranth and / or other dark green leaves, cassava leaves, etc.) ?
FCSVegGreSRf | Over the last 7 days,What was the main source of: Green leafy vegetables, such as: (spinach, broccoli, amaranth and / or other dark green leaves, cassava leaves, etc.) ?
FCSVegOth | Over the last 7 days, how many days did members of your household eat: Other Vegetables, such as: (Onion, tomatos, cucumber, radish, green beans, peas, etc) ?
HDDSVegOth | Yesterday, during the day and night, did members in your household eat: Other Vegetables, such as: (Onion, tomatos, cucumber, radish, green beans, peas, etc) ?
FCSVegOthSRf | Over the last 7 days,What was the main source of: Other Vegetables, such as: (Onion, tomatos, cucumber, radish, green beans, peas, etc) ?
FCSFruit | Over the last 7 days, how many days did members of your household eat: Fruits, such as:(banana, apple, lemon, mango, papaya, apricot, peach, etc.) ?
FCSFruitOrg | Over the last 7 days, how many days did members of your household eat: Orange fruits, such as: (mango, papaya, apricot, peach, etc.)
HDDSFruitOrg | Yesterday, during the day and night, did members in your household eat: Orange fruits, such as: (mango, papaya, apricot, peach, etc.)?
FCSFruitOrgSRf | Over the last 7 days,What was the main source of: Orange fruits, such as: (mango, papaya, apricot, peach, etc.) ?
FCSFruitOth | Over the last 7 days, how many days did members of your household eat: Other fruits, such as: (Pinapple, apple, lemon, etc.)?
HDDSFruitOth | Yesterday, during the day and night, did members in your household eat: Orange fruits, such as: (mango, papaya, apricot, peach, etc.)?
FCSFruitOthSRF | Over the last 7 days, how many days did members of your household eat: Other fruits, such as: (Pinapple, apple, lemon, etc.)?
FCSFat | Over the last 7 days,What was the main source of:  Fats and oils,such as: (vegetable oil, palm oil, shea butter, margarine, other fats /oil)?
HDDSFat | Yesterday, during the day and night, did members in your household eat: Fats and oils,such as: (vegetable oil, palm oil, shea butter, margarine, other fats /oil)?
FCSFatSRf | Over the last 7 days,What was the main source of:  Fats and oils,such as: (vegetable oil, palm oil, shea butter, margarine, other fats /oil)?
FCSSugar | Over the last 7 days,What was the main source of:  Sugar and sweets, such as: (sugar, honey, jam, cakes, candy, cookies, pastries, cakes and other sweet sugary drinks)?
HDDSSugar | Yesterday, during the day and night, did members in your household eat: Sugar and sweets, such as:(sugar, honey, jam, cakes, candy, cookies, pastries, cakes and other sweet sugary drinks)?
FCSSugarSRf | Over the last 7 days,What was the main source of: Sugar and sweets, such as: (sugar, honey, jam, cakes, candy, cookies, pastries, cakes and other sweet sugary drinks)?
FCSCond | Over the last 7 days, What was the main source of: Condiments/Spices: such as: (tea, coffee / cocoa, salt, garlic, spices, yeast / baking powder, tomato / sauce, meat or fish as a condiment, condiments including small amount of milk / tea coffee)?
HDDSCond | Yesterday, during the day and night, did members in your household eat: Condiments/Spices: such as: (tea, coffee / cocoa, salt, garlic, spices, yeast / baking powder, tomato / sauce, meat or fish as a condiment, condiments including small amount of milk / tea coffee)?
FCSCondSRf | Over the last 7 days,What was the main source of: Condiments/Spices: such as: (tea, coffee / cocoa, salt, garlic, spices, yeast / baking powder, tomato / sauce, meat or fish as a condiment, condiments including small amount of milk / tea coffee)?

Here is the standardized module in xlsform:

[RBDstandardized_questionnaireFCSHDDS](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/questionnaires/RBDstandardized_questionnaireFCSHDDS.xlsx) 


## Calculation of FCS & HDDS Indicators & CH 

### Example data set

Here is the example data set:

still need to create 

### SPSS Syntax

```{r, eval = F}

to be added

```

Here is the SPSS syntax file: 

not done yet

### R Syntax

```{r, eval = F}
library(haven)
library(labelled)
library(tidyverse)

#import dataset
dataFCSHDDSEng <- read_sav("dataFCSHDDSEng.sav")

#Calculate HHS 
dataFCSHDDSEng <- to_factor(dataFCSHDDSEng)


```



Here is the R syntax file: 

needs to be done

## References / Resources


<!--chapter:end:03-FCSHDDS.Rmd-->

# Livelihood Coping Strategies

## Standardized Module - Light - Coping for Food

Variable Name | Question Label | Answer Choices
---- | ------------- |------- 
LhCSIStress1 | During the past 30 days, did anyone in your household have to: *Sell household assets/goods  radio, furniture, refrigerator, television, jewelry etc..)* due to a lack of food or a lack of money to buy food? | 1) No, because I did not face a shortage of food 2) No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it 3) Yes 4) Not applicable
LhCSIStress2 | During the past 30 days, did anyone in your household have to: *sell more animals (non-productive) than usual* due to a lack of food or a lack of money to buy food? | 1) No, because I did not face a shortage of food 2) No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it 3) Yes 4) Not applicable
LhCSIStress3 | During the past 30 days, did anyone in your household have to: *spend savings* due to a lack of food or a lack of money to buy food? | 1) No, because I did not face a shortage of food 2) No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it 3) Yes 4) Not applicable
LhCSIStress4 | During the past 30 days, did anyone in your household have to: *Borrow money / food  from a formal lender / bank* due to a lack of food or a lack of money to buy food? | 1) No, because I did not face a shortage of food 2) No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it 3) Yes 4) Not applicable
LhCSICrisis1 | During the past 30 days, did anyone in your household have to: *Reduce non-food expenses on health (including drugs) and education * due to a lack of food or a lack of money to buy food? | 1) No, because I did not face a shortage of food 2) No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it 3) Yes 4) Not applicable
LhCSICrisis2 | During the past 30 days, did anyone in your household have to: *sell productive assets or means of transport (sewing machine, wheelbarrow, bicycle, car, etc..)* due to a lack of food or a lack of money to buy food? | 1) No, because I did not face a shortage of food 2) No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it 3) Yes 4) Not applicable
LhCSICrisis3 | During the past 30 days, did anyone in your household have to: *withdraw children from school* due to a lack of food or a lack of money to buy food? | 1) No, because I did not face a shortage of food 2) No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it 3) Yes 4) Not applicable
LhCSIEmergency1 | During the past 30 days, did anyone in your household have to: *sell house or land* due to a lack of food or a lack of money to buy food? | 1) No, because I did not face a shortage of food 2) No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it 3) Yes 4) Not applicable
LhCSIEmergency2 | During the past 30 days, did anyone in your household have to: *sell last female animals* due to a lack of food or a lack of money to buy food? | 1) No, because I did not face a shortage of food 2) No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it 3) Yes 4) Not applicable
LhCSIEmergency3 | During the past 30 days, did anyone in your household have to: *beg * due to a lack of food or a lack of money to buy food? | 1) No, because I did not face a shortage of food 2) No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it 3) Yes 4) Not applicable

### Standardized Module - data collection form

Here is the standardized module in xlsform: [RBDstandardized_questionnaireLHCS](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/questionnaires/RBDstandardized_questionnaireLHCS.xlsx) 

### Analysis 

#### Example data set
[dataLHCSLightEng](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/example_datasets/dataLHCSLightEng.sav) 

Here is the example data set:

#### SPSS Syntax

```{r, eval = F}
```

Here is the SPSS syntax file: 

still need to do


#### R Syntax

```{r, eval = F}
library(haven)
library(labelled)
library(tidyverse)

#import dataset
dataLHCSLightEng <- read_sav("dataLHCSLightEng.sav")

#Calculate HHS 
dataLHCSLightEng <- to_factor(dataLHCSLightEng)

#create a variable to specify if the household used any of the strategies by severity 
#stress
dataLHCSLightEng <- dataLHCSLightEng %>% mutate(stress_coping = case_when(
LhCSIStress1 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedstress",
LhCSIStress2 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedstress",
LhCSIStress3 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedstress",
LhCSIStress4 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedstress",
TRUE ~ "notusedstress"))
#Crisis
dataLHCSLightEng <- dataLHCSLightEng %>% mutate(crisis_coping = case_when(
LhCSICrisis1 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedcrisis",
LhCSICrisis2 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedcrisis",
LhCSICrisis3 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedcrisis",
TRUE ~ "notusedcrisis"))
#Emergency
dataLHCSLightEng <- dataLHCSLightEng %>% mutate(emergency_coping = case_when(
LhCSIEmergency1 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedemergency",
LhCSIEmergency2 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedemergency",
LhCSIEmergency3 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedemergency",
TRUE ~ "noutusedemergency"))

#calculate Max_coping_behaviour
dataLHCSLightEng <- dataLHCSLightEng %>% mutate(LhCSICat = case_when(
  emergency_coping == "usedemergency" ~ "EmergencyStrategies",
  crisis_coping == "usedcrisis" ~ "CrisisStrategies",
  stress_coping == "usedstress" ~ "StressStrategies",
  TRUE ~ "NoStrategies"))

#Generate table of proportion of households in CH LhCHS phases by Adm1 and Adm2 using weights
#Livelihood Coping Strategies 
LhHCSCat_table_wide <- dataLHCSLightEng %>% 
  drop_na(LhCSICat) %>%
  group_by(ADMIN1Name, ADMIN2Name) %>%
  count(LhCSICat, wt = WeightHH) %>%
  mutate(perc = 100 * n / sum(n)) %>%
  ungroup() %>% select(-n) %>%
  spread(key = LhCSICat, value = perc) %>% replace(., is.na(.), 0) %>% mutate_if(is.numeric, round, 1)
```


Here is the R syntax file: 
[RBDstandardized_RsyntaxLHCSIlight](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/syntax/RBDstandardized_RsyntaxLHCSIlight.R) 


## Standardized Module - Light - Essential Needs Assessment (ENA)

Variable Name | Question Label | Answer Choices
---- | ------------- |--------------- 
LhCSIStress1_EN | During the past 30 days, did anyone in your household have to: *Sell household assets/goods (radio, furniture, refrigerator, television, jewelry etc..)* because there was not enough resources (food, cash, else) to meet essential needs (e.g. adequate shelter, education services, health services, food, etc)? | 1) No, because I did not face a shortage of food 2) No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it 3) Yes 4) Not applicable
LhCSIStress1_EN_why | What was the MAIN REASON you or other members in your household adopted these coping strategies to access essential needs? | 1)To access food 2)To access education services/commodities (e.g. uniforms, books) 3)To access health services/medicines 4)To buy and repair cloths or shoes 5)To access adequate shelter 6)To access water/sanitation facilities 7)To access essential dwelling services (electricity, energy, waste disposal…) 8)To pay for existing debts 9)Other
LhCSIStress2_EN | During the past 30 days, did anyone in your household have to: *Purchase food on credit or borrow food* because there was not enough resources (food, cash, else) to meet essential needs (e.g. adequate shelter, education services, health services, food, etc)? | 1) No, because I did not face a shortage of food 2) No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it 3) Yes 4) Not applicable
LhCSIStress2_EN_why | What was the MAIN REASON you or other members in your household adopted these coping strategies to access essential needs? | 1)To access food 2)To access education services/commodities (e.g. uniforms, books) 3)To access health services/medicines 4)To buy and repair cloths or shoes 5)To access adequate shelter 6)To access water/sanitation facilities 7)To access essential dwelling services (electricity, energy, waste disposal…) 8)To pay for existing debts 9)Other
LhCSIStress3_EN | During the past 30 days, did anyone in your household have to: *spend savings* because there was not enough resources (food, cash, else) to meet essential needs (e.g. adequate shelter, education services, health services, food, etc)? | 1) No, because I did not face a shortage of food 2) No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it 3) Yes 4) Not applicable
LhCSIStress3_EN_why | What was the MAIN REASON you or other members in your household adopted these coping strategies to access essential needs? | 1)To access food 2)To access education services/commodities (e.g. uniforms, books) 3)To access health services/medicines 4)To buy and repair cloths or shoes 5)To access adequate shelter 6)To access water/sanitation facilities 7)To access essential dwelling services (electricity, energy, waste disposal…) 8)To pay for existing debts 9)Other
LhCSIStress4_EN | During the past 30 days, did anyone in your household have to: *Borrow money* because there was not enough resources (food, cash, else) to meet essential needs (e.g. adequate shelter, education services, health services, food, etc)? | 1) No, because I did not face a shortage of food 2) No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it 3) Yes 4) Not applicable
LhCSIStress4_EN_why | What was the MAIN REASON you or other members in your household adopted these coping strategies to access essential needs? | 1)To access food 2)To access education services/commodities (e.g. uniforms, books) 3)To access health services/medicines 4)To buy and repair cloths or shoes 5)To access adequate shelter 6)To access water/sanitation facilities 7)To access essential dwelling services (electricity, energy, waste disposal…) 8)To pay for existing debts 9)Other
LhCSICrisis1_EN | During the past 30 days, did anyone in your household have to: *sell productive assets or means of transport (sewing machine, wheelbarrow, bicycle, car, etc..)* because there was not enough resources (food, cash, else) to meet essential needs (e.g. adequate shelter, education services, health services, food, etc)? | 1) No, because I did not face a shortage of food 2) No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it 3) Yes 4) Not applicable
LhCSICrisis1_EN_why | What was the MAIN REASON you or other members in your household adopted these coping strategies to access essential needs? | 1)To access food 2)To access education services/commodities (e.g. uniforms, books) 3)To access health services/medicines 4)To buy and repair cloths or shoes 5)To access adequate shelter 6)To access water/sanitation facilities 7)To access essential dwelling services (electricity, energy, waste disposal…) 8)To pay for existing debts 9)Other
LhCSICrisis2_EN | During the past 30 days, did anyone in your household have to: *consume seed stocks that were to be held/saved for the next season* because there was not enough resources (food, cash, else) to meet essential needs (e.g. adequate shelter, education services, health services, food, etc)? | 1) No, because I did not face a shortage of food 2) No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it 3) Yes 4) Not applicable
LhCSICrisis2_EN_why | What was the MAIN REASON you or other members in your household adopted these coping strategies to access essential needs? | 1)To access food 2)To access education services/commodities (e.g. uniforms, books) 3)To access health services/medicines 4)To buy and repair cloths or shoes 5)To access adequate shelter 6)To access water/sanitation facilities 7)To access essential dwelling services (electricity, energy, waste disposal…) 8)To pay for existing debts 9)Other
LhCSICrisis3_EN | During the past 30 days, did anyone in your household have to: *Withdraw children from school* because there was not enough resources (food, cash, else) to meet essential needs (e.g. adequate shelter, education services, health services, food, etc)? | 1) No, because I did not face a shortage of food 2) No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it 3) Yes 4) Not applicable
LhCSICrisis3_EN_why | What was the MAIN REASON you or other members in your household adopted these coping strategies to access essential needs? | 1)To access food 2)To access education services/commodities (e.g. uniforms, books) 3)To access health services/medicines 4)To buy and repair cloths or shoes 5)To access adequate shelter 6)To access water/sanitation facilities 7)To access essential dwelling services (electricity, energy, waste disposal…) 8)To pay for existing debts 9)Other
LhCSIEmergency1_EN | During the past 30 days, did anyone in your household have to: *Sell house or land because there was not enough resources (food, cash, else)*  because there was not enough resources (food, cash, else) to meet essential needs (e.g. adequate shelter, education services, health services, food, etc)? | 1) No, because I did not face a shortage of food 2) No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it 3) Yes 4) Not applicable
LhCSIEmergency1_EN_why | What was the MAIN REASON you or other members in your household adopted these coping strategies to access essential needs? | 1)To access food 2)To access education services/commodities (e.g. uniforms, books) 3)To access health services/medicines 4)To buy and repair cloths or shoes 5)To access adequate shelter 6)To access water/sanitation facilities 7)To access essential dwelling services (electricity, energy, waste disposal…) 8)To pay for existing debts 9)Other
LhCSIEmergency2_EN | During the past 30 days, did anyone in your household have to: *Beg* because there was not enough resources (food, cash, else) to meet essential needs (e.g. adequate shelter, education services, health services, food, etc)? | 1) No, because I did not face a shortage of food 2) No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it 3) Yes 4) Not applicable
LhCSIEmergency2_EN_why | What was the MAIN REASON you or other members in your household adopted these coping strategies to access essential needs? | 1)To access food 2)To access education services/commodities (e.g. uniforms, books) 3)To access health services/medicines 4)To buy and repair cloths or shoes 5)To access adequate shelter 6)To access water/sanitation facilities 7)To access essential dwelling services (electricity, energy, waste disposal…) 8)To pay for existing debts 9)Other
LhCSIEmergency3_EN| During the past 30 days, did anyone in your household have to: *Sell last female animals* because there was not enough resources (food, cash, else) to meet essential needs (e.g. adequate shelter, education services, health services, food, etc)? | 1) No, because I did not face a shortage of food 2) No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it 3) Yes 4) Not applicable
LhCSIEmergency3_EN_why | What was the MAIN REASON you or other members in your household adopted these coping strategies to access essential needs? | 1)To access food 2)To access education services/commodities (e.g. uniforms, books) 3)To access health services/medicines 4)To buy and repair cloths or shoes 5)To access adequate shelter 6)To access water/sanitation facilities 7)To access essential dwelling services (electricity, energy, waste disposal…) 8)To pay for existing debts 9)Other

### Standardized Module - data collection form

Here is the standardized module in xlsform: [RBDstandardized_questionnaireLHCS](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/questionnaires/RBDstandardized_questionnaireLHCS.xlsx) 

### Analysis 

#### Example data set
[dataLHCSLightEng](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/example_datasets/dataLHCSLightEng.sav) 

Here is the example data set:

#### SPSS Syntax

```{r, eval = F}
```

Here is the SPSS syntax file: 

still need to do


#### R Syntax

```{r, eval = F}
library(haven)
library(labelled)
library(tidyverse)

#import dataset
dataLHCSENAEng <- read_sav("dataLHCSENAEng.sav")

#Calculate HHS 
dataLHCSENAEng <- to_factor(dataLHCSENAEng)

#create a variable to specify if the household used any of the strategies by severity 
#stress
dataLHCSENAEng <- dataLHCSENAEng %>% mutate(stress_coping = case_when(
LhCSIStress1 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedstress",
LhCSIStress2 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedstress",
LhCSIStress3 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedstress",
LhCSIStress4 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedstress",
TRUE ~ "notusedstress"))
#Crisis
dataLHCSLightEng <- dataLHCSLightEng %>% mutate(crisis_coping = case_when(
LhCSICrisis1 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedcrisis",
LhCSICrisis2 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedcrisis",
LhCSICrisis3 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedcrisis",
TRUE ~ "notusedcrisis"))
#Emergency
dataLHCSLightEng <- dataLHCSLightEng %>% mutate(emergency_coping = case_when(
LhCSIEmergency1 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedemergency",
LhCSIEmergency2 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedemergency",
LhCSIEmergency3 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "usedemergency",
TRUE ~ "noutusedemergency"))

#calculate Max_coping_behaviour
dataLHCSLightEng <- dataLHCSLightEng %>% mutate(LhCSICat = case_when(
  emergency_coping == "usedemergency" ~ "EmergencyStrategies",
  crisis_coping == "usedcrisis" ~ "CrisisStrategies",
  stress_coping == "usedstress" ~ "StressStrategies",
  TRUE ~ "NoStrategies"))

#Generate table of proportion of households in CH LhCHS phases by Adm1 and Adm2 using weights
#Livelihood Coping Strategies 
LhHCSCat_table_wide <- dataLHCSLightEng %>% 
  drop_na(LhCSICat) %>%
  group_by(ADMIN1Name, ADMIN2Name) %>%
  count(LhCSICat, wt = WeightHH) %>%
  mutate(perc = 100 * n / sum(n)) %>%
  ungroup() %>% select(-n) %>%
  spread(key = LhCSICat, value = perc) %>% replace(., is.na(.), 0) %>% mutate_if(is.numeric, round, 1)
```


Here is the R syntax file: 
[RBDstandardized_RsyntaxLHCSIlight](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/syntax/RBDstandardized_RsyntaxLHCSIlight.R) 


<!--chapter:end:04-LHCS.Rmd-->

# Gecodes

Using names as identifiers can easily lead to confusion over spelling and transliterations as well as alternative and duplicate names.

As unique identifiers, P-Codes overcome the challenges of linking datasets and confirming locations.

We recommend using Gecodes and Names from OCHA's Common Operational Datasets: 
https://data.humdata.org/dashboards/cod?ext_geodata=1&q=&ext_page_size=25

## A brief example of the problem:

The geographic variables of the *choices* section of the  xlsform typically look like this

list_name | name | label
---- | --- |---- 
cod_lga | 1 | ASKIRA/UBA
cod_lga | 2 | BAYO
cod_lga | 3 | CHIBOK
cod_lga | 4 | KALABAGE

The geographic variables in corresponding GIS dataset look like this:

Shape | admin2Pcode | admin2RefName
---- | --- |---- 
Polygon |NG008002 | Askira/Uba
Polygon |NG008004 | Bayo
Polygon |NG008006 | Chibok
Polygon |NG008015 | Kala/Balge

Unfortunately, if we tried to merge the dataset and the GIS file by the name of the area, we would first have to fix the capitalization and punctuation of the names.

To avoid unnecessary work, we recommend putting in the *Pcode* in the name column of the xlsform:

list_name | name | label
---- | --- |---- 
cod_lga | NG008002 | ASKIRA/UBA
cod_lga | NG008004 | BAYO
cod_lga | NG008006 | CHIBOK
cod_lga | NG008015 | KALABAGE

With a common, simple variable to merge both datasets, mapping and data processing will be much easier!

## Where to find Pcodes

Geographic data can be found https://data.humdata.org/dashboards/cod?ext_geodata=1&q=&ext_page_size=25 and searching for the appropriate country.

We recommend using the codes from the *admin1Pcode* and *admin2Pcode* columns in the *name* column of the xlsform.  Using the *admin1RefName* and *admin2RefName*

## Example data set of the data collection sheet for Nigeria:
[xlsformwithgeocodesexample](https://github.com/WFP-VAM/RBD_FS_CH_guide_EN/blob/master/questionnaires/xlsformwithgeocodesexample.xlsx) 



## 2 caveats

1. In some countries, the boundaries in the CODS might not correspond with the latest boundaries used by the country.  In this case, special development of a GIS boundary file and codes will need to be developed.

2. Inserting the relevant pcodes in the *name* column will probably require some *vlookup* .  This might be relatively straightforward when only replacing codes at the admin1 and admin2 level, however things might get more complicated and difficult if cascading columns and the *choice_filter* option is used at the adm3, adm4 level. 




<!--chapter:end:05-Geocodes.Rmd-->

# Example of Data Quality Report

<!--chapter:end:06-DataQuality.Rmd-->

