---
title: "Example Data Quality Monitoring Report"
author: "WFP "
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F, warning = F,  fig.width=10, fig.height=10)
library(tidyverse)
library(haven)
library(labelled)
library(plotly)

#set working directory 
setwd("C:/exampledataquality")

#import dataset and call it "dataset"
dataset <- read_sav("1_RawData/exampledataEnglish_raw.sav")

#convert into labels for all variables
dataset <- to_factor(dataset)

#change back to value to prevent showing operators name
dataset$EnumName <- to_labelled(dataset$EnumName )


#Survey Progress
countsadm1table <- dataset %>% group_by(ADMIN1Name) %>%  count()
countsadm1adm2table <- dataset %>% group_by(ADMIN1Name, ADMIN2Name) %>%  count() 
countsenumtable <- dataset %>% group_by(ADMIN2Name, EnumName) %>%  count() %>% arrange(desc(`n`))

adm1_graph <- countsadm1table %>% ggplot(aes(x = ADMIN1Name, y = n)) +geom_bar(stat = "identity", fill = "steel blue") +theme_minimal()  +labs(x = "", y = "")
adm2_graph <- countsadm1adm2table %>% ggplot(aes(x = ADMIN2Name, y = n)) +geom_bar(stat = "identity", fill = "steel blue") +theme_minimal() +facet_wrap(. ~ ADMIN1Name, scales = "free") +theme(axis.text.x = element_text(angle = 90)) +labs(x = "", y = "")
adm1enum_graph <- countsenumtable %>% ggplot(aes(x = as.factor(EnumName), y = n)) +geom_bar(stat = "identity", fill = "steel blue") +theme_minimal()  +facet_wrap(. ~ ADMIN2Name, scales = "free_x") +labs(x = "", y = "") +theme(axis.text.x=element_blank())

#Food Consumption Score
#calculate FCS
dataset <- dataset %>% mutate(FCS = (2 * FCSStap) + (3 * FCSPulse)+ (4*FCSPr) +FCSVeg  +FCSFruit +(4*FCSDairy) + (0.5*FCSFat) + (0.5*FCSSugar))
#create FCG groups based on 21/25 or 28/42 thresholds
dataset <- dataset %>% mutate(
  FCSCat21 = case_when(
    FCS <= 21 ~ "Poor", between(FCS, 21.5, 35) ~ "Borderline", FCS > 35 ~ "Acceptable"),
  FCSCat28 = case_when(
    FCS <= 28 ~ "Poor", between(FCS, 28.5, 42) ~ "Borderline", FCS > 42 ~ "Acceptable"))
#fcg colors
fcg_colors = c("Acceptable" = "#27AE60","Borderline" = "#F1C40F","Poor" = "#C0392B")
#FCG by Adm2
FCGadm2table <- dataset %>% group_by(ADMIN1Name, ADMIN2Name) %>%  count(FCSCat28) %>% mutate(perc = 100 * n / sum(n)) %>% ungroup() %>% select(-n) %>% mutate_if(is.numeric, round, 1)
FCGadm2_barplot <- FCGadm2table %>% ggplot(aes(x=ADMIN2Name, y = perc, fill = FCSCat28)) +geom_bar(position="fill", stat = "identity") +theme_minimal() +theme(axis.text.x = element_text(angle = 90)) +scale_fill_manual(values=fcg_colors) + facet_wrap(. ~ ADMIN1Name, scales = "free_x")  +labs(x = "", y = "") +scale_y_continuous(labels = scales::percent)
#FCG by Adm2 and EnumName
FCGEnumNametable <- dataset %>% group_by(EnumName, ADMIN1Name, ADMIN2Name) %>%  count(FCSCat28) %>% mutate(perc = 100 * n / sum(n)) %>% ungroup() %>% select(-n) %>% mutate_if(is.numeric, round, 1)
FCGadm2EnumName_barplot <- FCGEnumNametable %>% ggplot(aes(x=as.factor(EnumName), y = perc, fill = FCSCat28)) +geom_bar(position="fill", stat = "identity") +theme_minimal() +theme(axis.text.x = element_text(angle = 90)) +scale_fill_manual(values=fcg_colors) + facet_wrap(ADMIN1Name ~ ADMIN2Name, scales = "free") +labs(x = "", y = "") +theme(axis.text.x=element_blank()) +scale_y_continuous(labels = scales::percent)

#Household Dietary Diversity Score
#calculate HDDS first by creating the 12 groups based on the 16 questions
dataset <- dataset %>% mutate(
  HDDSStapCer = case_when(HDDSStapCer == "Yes" ~ 1, TRUE ~ 0),
  HDDSStapRoot = case_when(HDDSStapRoot  == "Yes" ~ 1, TRUE ~ 0),
  HDDSVeg = case_when(HDDSVegOrg  == "Yes" | HDDSVegGre == "Yes" | HDDSVegOth == "Yes" ~ 1, TRUE ~ 0),
  HDDSFruit = case_when(HDDSFruitOrg == "Yes" | HDDSFruitOth == "Yes" ~ 1, TRUE ~ 0),
  HDDSPrMeat = case_when(HDDSPrMeatF == "Yes" | HDDSPrMeatO == "Yes" ~ 1, TRUE ~ 0),
  HDDSPrEgg = case_when(HDDSPrEgg  == "Yes" ~ 1, TRUE ~ 0),
  HDDSPrFish = case_when(HDDSPrFish == "Yes" ~ 1, TRUE ~ 0),
  HDDSPulse = case_when(HDDSPulse == "Yes" ~ 1, TRUE ~ 0),
  HDDSDairy = case_when(HDDSDairy == "Yes" ~ 1, TRUE ~ 0),
  HDDSFat = case_when(HDDSFat == "Yes" ~ 1, TRUE ~ 0),
  HDDSSugar = case_when(HDDSSugar == "Yes" ~ 1, TRUE ~ 0),
  HDDSCond = case_when(HDDSCond == "Yes"~ 1, TRUE ~ 0))
#Calculate HDDS and Cadre Harmonise Phases
dataset <- dataset %>% mutate(HDDS = HDDSStapCer +HDDSStapRoot +HDDSVeg +HDDSFruit +HDDSPrMeat +HDDSPrEgg +HDDSPrFish +HDDSPulse +HDDSDairy +HDDSFat +HDDSSugar +HDDSCond)
#Calucate Cadre Harmonise HDDS phasing categories
dataset<- dataset %>%  mutate(HDDS_CH = case_when(
  HDDS >= 5 ~ "Phase1",
  HDDS == 4 ~ "Phase2",
  HDDS == 3 ~ "Phase3",
  HDDS == 2 ~ "Phase4",
  HDDS < 2 ~ "Phase5"))
#CH colors
CH_colors = c("Phase1" = "#c6ffc7","Phase2" = "#ffe718","Phase3" = "#e88400","Phase4" = "#e02d00", "Phase5" = "#5e0803")
#HDDS phases by Adm2
HDDSphaseadm2table <- dataset %>% group_by(ADMIN1Name, ADMIN2Name) %>%  count(HDDS_CH) %>% mutate(perc = 100 * n / sum(n)) %>% ungroup() %>% select(-n) %>% mutate_if(is.numeric, round, 1)
HDDSadm2_barplot <- HDDSphaseadm2table %>% ggplot(aes(x=ADMIN2Name, y = perc, fill = HDDS_CH)) +geom_bar(position="fill", stat = "identity") +theme_minimal() +theme(axis.text.x = element_text(angle = 90)) +scale_fill_manual(values=CH_colors) + facet_wrap(. ~ ADMIN1Name, scales = "free_x") +labs(x = "", y = "") +scale_y_continuous(labels = scales::percent)
#FCG by Adm2 and EnumName
HDDSadm2EnumNametable <- dataset %>% group_by(EnumName,  ADMIN1Name, ADMIN2Name) %>%  count(HDDS_CH) %>% mutate(perc = 100 * n / sum(n)) %>% ungroup() %>% select(-n) %>% mutate_if(is.numeric, round, 1)
HDDSadm2EnumName_barplot <- HDDSadm2EnumNametable %>% ggplot(aes(x=as.factor(EnumName), y = perc, fill = HDDS_CH)) +geom_bar(position="fill", stat = "identity") +theme_minimal() +theme(axis.text.x = element_text(angle = 90)) +scale_fill_manual(values=CH_colors) + facet_wrap(ADMIN1Name ~ ADMIN2Name, scales = "free") +labs(x = "", y = "") +theme(axis.text.x=element_blank()) +scale_y_continuous(labels = scales::percent)



#reduced Coping Strategy Index
#calculate rCSI Score
dataset <- dataset %>% mutate(rCSI = rCSILessQlty  + (2 * rCSIBorrow) + rCSIMealSize + (3 * rCSIMealAdult) + rCSIMealNb)
var_label(dataset$rCSI) <- "rCSI"
#Create rCSI Cadre Harmonise Categories
dataset<- dataset %>% mutate(rCSI_CH = case_when(
  rCSI <= 3 ~ "Phase1",
  between(rCSI,4,18) ~ "Phase2",
  rCSI >= 19 ~ "Phase3"))
#rCSI colors
rCSI_colors = c("Phase1" = "#c6ffc7","Phase2" = "#ffe718","Phase3" = "#e88400")
#FCG by Adm2
rCSIphaseadm2table <- dataset %>% group_by(ADMIN1Name, ADMIN2Name) %>%  count(rCSI_CH) %>% mutate(perc = 100 * n / sum(n)) %>% ungroup() %>% select(-n) %>% mutate_if(is.numeric, round, 1)
rCSIadm2_barplot <- rCSIphaseadm2table %>% ggplot(aes(x=ADMIN2Name, y = perc, fill = rCSI_CH)) +geom_bar(position="fill", stat = "identity") +theme_minimal() +theme(axis.text.x = element_text(angle = 90)) +scale_fill_manual(values=rCSI_colors) + facet_wrap(. ~ ADMIN1Name, scales = "free_x") +labs(x = "", y = "") +scale_y_continuous(labels = scales::percent)
#FCG by Adm2 and EnumName
rCSIEnumNametable <- dataset %>% group_by(EnumName, ADMIN1Name, ADMIN2Name) %>%  count(rCSI_CH) %>% mutate(perc = 100 * n / sum(n)) %>% ungroup() %>% select(-n) %>% mutate_if(is.numeric, round, 1)
rCSIadm2EnumName_barplot <- rCSIEnumNametable %>% ggplot(aes(x=as.factor(EnumName), y = perc, fill = rCSI_CH)) +geom_bar(position="fill", stat = "identity") +theme_minimal() +theme(axis.text.x = element_text(angle = 90)) +scale_fill_manual(values=rCSI_colors) + facet_wrap(ADMIN1Name ~ ADMIN2Name, scales = "free") +labs(x = "", y = "") +theme(axis.text.x=element_blank()) +scale_y_continuous(labels = scales::percent)

#HHS Score 
#Recode HHS questions into new variables with score
dataset <- dataset %>% mutate(HHhSNoFood_FR_r = case_when(
  HHhSNoFood_FR == "Rarely (1–2 times)" ~ 1,
  HHhSNoFood_FR == "Sometimes (3–10 times)" ~ 1,
  HHhSNoFood_FR == "Often (more than 10 times)" ~ 2,
  TRUE ~ 0),
  HHhSBedHung_FR_r = case_when(
    HHhSBedHung_FR == "Rarely (1–2 times)" ~ 1,
    HHhSBedHung_FR == "Sometimes (3–10 times)" ~ 1,
    HHhSBedHung_FR == "Often (more than 10 times)" ~ 2,
    TRUE ~ 0),
  HHhSNotEat_FR_r = case_when(
    HHhSNotEat_FR == "Rarely (1–2 times)" ~ 1,
    HHhSNotEat_FR == "Sometimes (3–10 times)" ~ 1,
    HHhSNotEat_FR == "Often (more than 10 times)" ~ 2,
    TRUE ~ 0))
# Calculate HHhS score
dataset <- dataset %>% mutate(HHhS = HHhSNoFood_FR_r + HHhSBedHung_FR_r + HHhSNotEat_FR_r)
#Convert HH Scores to CH phases
dataset <- dataset %>% mutate(HHhS_CH = case_when(
  HHhS == 0 ~ "Phase1",
  HHhS == 1 ~ "Phase2",
  HHhS %in% c(2,3) ~ "Phase3",
  HHhS == 4 ~ "Phase4",
  HHhS >= 5 ~ "Phase5"))
#CH colors
CH_colors = c("Phase1" = "#c6ffc7","Phase2" = "#ffe718","Phase3" = "#e88400","Phase4" = "#e02d00", "Phase5" = "#5e0803")
#HHS by Adm2
HHSphaseadm2table <- dataset %>% group_by(ADMIN1Name, ADMIN2Name) %>%  count(HHhS_CH) %>% mutate(perc = 100 * n / sum(n)) %>% ungroup() %>% select(-n) %>% mutate_if(is.numeric, round, 1)
HHSadm2_barplot <- HHSphaseadm2table %>% ggplot(aes(x=ADMIN2Name, y = perc, fill = HHhS_CH)) +geom_bar(position="fill", stat = "identity") +theme_minimal() +theme(axis.text.x = element_text(angle = 90)) +scale_fill_manual(values=CH_colors) + facet_wrap(. ~ ADMIN1Name, scales = "free_x") +labs(x = "", y = "") +scale_y_continuous(labels = scales::percent)
#FCG by Adm2 and EnumName
HHSadm2EnumNametable <- dataset %>% group_by(EnumName,  ADMIN1Name, ADMIN2Name) %>%  count(HHhS_CH) %>% mutate(perc = 100 * n / sum(n)) %>% ungroup() %>% select(-n) %>% mutate_if(is.numeric, round, 1)
HHSadm2EnumName_barplot <- HHSadm2EnumNametable %>% ggplot(aes(x=as.factor(EnumName), y = perc, fill = HHhS_CH)) +geom_bar(position="fill", stat = "identity") +theme_minimal() +theme(axis.text.x = element_text(angle = 90)) +scale_fill_manual(values=CH_colors) + facet_wrap(ADMIN1Name ~ ADMIN2Name, scales = "free") +labs(x = "", y = "") +theme(axis.text.x=element_blank()) +scale_y_continuous(labels = scales::percent)


#Livelihood Coping Strategies (CARI)
#create a variable to specify if the household used any of the strategies by severity
#stress
dataset <- dataset %>% mutate(stress_coping = case_when(
  LhCSIStress1 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "Yes",
  LhCSIStress2 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "Yes",
  LhCSIStress3 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "Yes",
  LhCSIStress4 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "Yes",
  TRUE ~ "No"))
#Crisis
dataset <- dataset %>% mutate(crisis_coping = case_when(
  LhCSICrisis1 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "Yes",
  LhCSICrisis2 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "Yes",
  LhCSICrisis3 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "Yes",
  TRUE ~ "No"))
#Emergency
dataset <- dataset %>% mutate(emergency_coping = case_when(
  LhCSIEmergency1 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "Yes",
  LhCSIEmergency2 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "Yes",
  LhCSIEmergency3 %in% c("Yes","No; because I already sold those assets or did this activity in the last 12 months and cannot continue to do it") ~ "Yes",
  TRUE ~ "No"))
#calculate Max_coping_behaviour
dataset <- dataset %>% mutate(LhCSICat = case_when(
  emergency_coping == "Yes" ~ "EmergencyStrategies",
  crisis_coping == "Yes" ~ "CrisisStrategies",
  stress_coping == "Yes" ~ "StressStrategies",
  TRUE ~ "NoStrategies"))
#CH colors
LHCS_colors = c("NoStrategies" = "#c6ffc7","StressStrategies" = "#ffe718","CrisisStrategies" = "#e88400","EmergencyStrategies" = "#e02d00")
#change ordering
dataset <- dataset %>%
  mutate(LhCSICat = fct_relevel(LhCSICat, c("NoStrategies", "StressStrategies", "CrisisStrategies", "EmergencyStrategies")))
#LHCS by Adm2
LHCSadm2table <- dataset %>% group_by(ADMIN1Name, ADMIN2Name) %>%  count(LhCSICat) %>% mutate(perc = 100 * n / sum(n)) %>% ungroup() %>% select(-n) %>% mutate_if(is.numeric, round, 1)
LHCSadm2_barplot <- LHCSadm2table %>% ggplot(aes(x=ADMIN2Name, y = perc, fill = LhCSICat)) +geom_bar(position="fill", stat = "identity") +theme_minimal() +theme(axis.text.x = element_text(angle = 90)) +scale_fill_manual(values=LHCS_colors) + facet_wrap(. ~ ADMIN1Name, scales = "free_x") +labs(x = "", y = "") +scale_y_continuous(labels = scales::percent)
#LHCS by Adm2 and EnumName
LHCSadm2EnumNametable <- dataset %>% group_by(EnumName,  ADMIN1Name, ADMIN2Name) %>%  count(LhCSICat) %>% mutate(perc = 100 * n / sum(n)) %>% ungroup() %>% select(-n) %>% mutate_if(is.numeric, round, 1)
LHCSadm2EnumName_barplot <- LHCSadm2EnumNametable %>% ggplot(aes(x=as.factor(EnumName), y = perc, fill = LhCSICat)) +geom_bar(position="fill", stat = "identity") +theme_minimal() +theme(axis.text.x = element_text(angle = 90)) +scale_fill_manual(values=LHCS_colors) + facet_wrap(ADMIN1Name ~ ADMIN2Name, scales = "free") +labs(x = "", y = "") +theme(axis.text.x=element_blank()) +scale_y_continuous(labels = scales::percent)

```

## Survey Progress

### Survey Counts by Admin1

```{r}
ggplotly(adm1_graph)
```

### Survey Counts by Admin2

```{r}
ggplotly(adm2_graph)
```

### Survey Counts by Enumerator Code and Adm2

```{r}
ggplotly(adm1enum_graph)
```


## Food Consumption Score (FCS) / Food Consumption Groups (FCG)

### Food Consumption Groups by Admin2 

```{r}
ggplotly(FCGadm2_barplot)
```

### Food Consumption Groups by Admin2 and Enumerator

```{r}
ggplotly(FCGadm2EnumName_barplot)
```


## Household Dietary Diversity Score (HDDS)

### HDDS phases by Admin2 

```{r}
ggplotly(HDDSadm2_barplot)
```

### HDDS phases by Admin2 and Enumerator

```{r}
ggplotly(HDDSadm2EnumName_barplot)
```

## reduced Coping Strategy Index (rCSI)

### rCSI (CH phases) by Admin2 

```{r}
ggplotly(rCSIadm2_barplot)
```

### rCSI (CH phases) by Admin2 and Enumerator

```{r}
ggplotly(rCSIadm2EnumName_barplot)
```


## Household Hunger Scale (HHS)

### HHS (CH phases) by Admin2 

```{r}
ggplotly(HHSadm2_barplot)
```

### HHS (CH phases) by Admin2 and Enumerator

```{r}
ggplotly(HHSadm2EnumName_barplot)
```


## Livelihood Coping Strategies (LhCS)

### Livelihood Coping Strategies by Admin2 

```{r}
ggplotly(LHCSadm2_barplot)
```

### Livelihood Coping Strategies by Admin2 and Enumerator

```{r}
ggplotly(LHCSadm2EnumName_barplot)
```











